---
title: "Assignment7"
author: "yixiao"
date: "2019/12/2"
output: html_document
---

## Part I

#Upload data
```{r}
data <- read.csv("online.data.csv")
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)

# load the library for plot
# install.packages("ggplot2")
library(ggplot2)
library(reshape2)
hdata <- melt(data[,1:6] ,  id.vars = 'id', variable.name = 'score')
ggplot(hdata, aes(value)) + geom_histogram(color="darkblue", fill="lightblue") + facet_grid(score ~ .)

#Then visualize the relationships between variables

pairs(~post.test.score+pre.test.score+messages+forum.posts+av.assignment.score+level.up,data=data, 
   main="Scatterplot Matrix")

#Try to capture an intution about the data and the relationships

# Intuition: based on the plots, av.assignment.score could be a good indicator for predicting level.up

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

# install.packages("rpart")
library(rpart)
# rp <- rpart(level.up ~ post.test.score + messages + av.assignment.score, data = data, method="class")
rp <- rpart(level.up ~ post.test.score + messages, data = data, method="class")

#Plot and generate a CP table for your tree 

printcp(rp)

#Generate a probability value that represents the probability that a student levels up based your classification tree 

data$pred <- predict(rp, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}

library(ROCR)

#Plot the curve
pred.detail <- prediction(data$pred, data$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR


```

```{r}
#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

rp2 <- rpart(level.up ~ pre.test.score + forum.posts, data = data, method="class")
printcp(rp2)
data$pred2 <- predict(rp2, type = "prob")[,2]
#Plot the curve
pred2.detail <- prediction(data$pred2, data$level.up) 
plot(performance(pred2.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)
#Calculate the Area Under the Curve
unlist(slot(performance(pred2.detail,"auc"), "y.values"))

# model rp is better than model rp2, because the auc value of rp is higher
```


## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

threshold.pred1 <- 0.1

#Now generate three diagnostics:

prob <- as.data.frame(predict(rp, newdata = data, type = "p"))
res <- ifelse(prob$yes < threshold.pred1, "yes", "no")
cm = as.matrix(table(Label = data$level.up, Predicted = res))

data$accuracy.model1 <- sum(diag(cm))/sum(cm)

data$precision.model1 <- cm[2,2]/sum(cm[2,])

data$recall.model1 <- cm[2,2]/sum(cm[,2])

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons


#Convert to matrix


#Calculate kappa


#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

```

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.

