---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
install.packages("readr")
install.packages("ggplot2")
install.packages("ROCR")
```

```{r}
library(readr)
s <- read_csv("online.data.csv")
```
remove id 
```{r}
s1 <- dplyr::select(s, -1)
s1$level.up <- ifelse(s1$level.up=="yes", 1, 0)
```
wide to long format
```{r}
library(ggplot2)
s2 <- tidyr::gather(s1, "variable", "value", 1:6)
```

#Visualization 

##Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
```{r}
#Rplot1
ggplot(s2, aes(value))+
   geom_histogram(data = s2[s2$variable=="post.test.score",], binwidth = 0.01)+
   geom_histogram(data = s2[s2$variable=="pre.test.score",], binwidth = 0.01)+
   geom_histogram(data = s2[s2$variable=="messages",], binwidth = 5)+
   geom_histogram(data = s2[s2$variable=="forum.posts",], binwidth = 1)+
   geom_histogram(data = s2[s2$variable=="av.assignment.score",], binwidth = 0.01)+
   geom_histogram(data = s2[s2$variable=="level.up",], binwidth = 0.1)+
   facet_wrap(~variable, scale="free")
```
#Then visualize the relationships between variables
```{r}
#Rplot2
plot(s1)
```
#Corelation plot
```{r}
#Rplot3
COR <- cor(s1) 
library(corrplot)
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
```
Messages has quite a strng relation ith post.test.score and even the others. while forum.posts dont have such strong corelations


#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
rp <- rpart(level.up ~ messages + post.test.score + av.assignment.score, method="class", data=s1, control = rpart.control(minsplit=1, minbucket=1, cp=0.01))
```

```{r}
#Plot and generate a CP table for your tree 
printcp(rp)
post(rp, file = "tree.ps", title = "predicting levels up")
```

```{r}
#Generate a probability value that represents the probability that a student levels up based your classification tree 

s1$pred <- predict(rp, type = "prob")[,2]
#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```

```{r}
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve - Rplot 4
pred.detail <- prediction(s1$pred, s1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)
```

```{r}
#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))
#Unlist liberates the AUC value from the "performance" object created by ROCR
```

```{r}
#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
rp2 <- rpart(level.up ~ pre.test.score + forum.posts, method="class", data=s1, control = rpart.control(minsplit=1, minbucket=1))
printcp(rp2)
post(rp2, file = "tree2.ps", title = "predicting levels up")

s1$pred2 <- predict(rp2, type = "prob")[,2]
pred.detail2 <- prediction(s1$pred2, s1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)
#Rplot5

unlist(slot(performance(pred.detail2,"auc"), "y.values"))

##Since AUC of c.tree1 is 1 which is greater than AUC of c.tree2, 0.818, c.tree1 is a better model to predict "levels up" than c.tree2.
```




#Thresholds

```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

s1$threshold.pred1 <- ifelse(s1$pred2>0.7, 1, 0) 
#make a variable and choose may be say .7 as your limit value
```

#Now generate three diagnostics:

#Accuracy Model
```{r}
s3 <- dplyr::count(s1, level.up, threshold.pred1)
```

```{r}
s3$type <- ifelse(s3$level.up==1 & s3$threshold.pred1==1, "TP", ifelse(s3$level.up==1 & s3$threshold.pred1==0, "FN", ifelse(s3$level.up==0 & s3$threshold.pred1==1, "FP", "TN")))
```

```{r}
s4 <- data.frame((s3[s3$type=="TP",]$n + s3[s3$type=="TN",]$n) / sum(s3$n))
names(s4) <- c("accuracy.model1")
```

#precision model
```{r}
s4$precision.model1 <- s3[s3$type=="TP",]$n / (s3[s3$type=="TP",]$n + s3[s3$type=="FP",]$n)
```

#recall model
```{r}
s4$recall.model1 <- s3[s3$type=="TP",]$n / (s3[s3$type=="TP",]$n + s3[s3$type=="TN",]$n) 
```


```{r}
#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(s1$level.up, s1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
s4$kappa.model1 <- kappa(matrix1, exact = TRUE)/kappa(matrix1)


```

#Now choose a different threshold value and repeat these diagnostics. Threshold now is 0.6. Before threshold was 0.7


```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

s1$threshold.pred1 <- ifelse(s1$pred2>0.6, 1, 0) 
#make a variable and choose may be say .7 as your limit value
```

#Now generate three diagnostics:

#Accuracy Model
```{r}
s3 <- dplyr::count(s1, level.up, threshold.pred1)
```

```{r}
s3$type <- ifelse(s3$level.up==1 & s3$threshold.pred1==1, "TP", ifelse(s3$level.up==1 & s3$threshold.pred1==0, "FN", ifelse(s3$level.up==0 & s3$threshold.pred1==1, "FP", "TN")))
```

```{r}
s5 <- data.frame((s3[s3$type=="TP",]$n + s3[s3$type=="TN",]$n) / sum(s3$n))
names(s4) <- c("accuracy.model1")
```

#precision model
```{r}
s5$precision.model1 <- s3[s3$type=="TP",]$n / (s3[s3$type=="TP",]$n + s3[s3$type=="FP",]$n)
```

#recall model
```{r}
s5$recall.model1 <- s3[s3$type=="TP",]$n / (s3[s3$type=="TP",]$n + s3[s3$type=="TN",]$n) 
```


```{r}
#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(s1$level.up, s1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
s5$kappa.model1 <- kappa(matrix1, exact = TRUE)/kappa(matrix1)
```

threshold value of 0.6 looks definitely more better than 0.7 because of the final results