---
title: "Assignment 7 - Answers"
author: "Linh Doan"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
M1 <- read.csv("online.data.csv", sep = ",", header = TRUE)
#change level.up into numberic value No=0 and Y=1
M1$level.up.numeric <- ifelse (M1$level.up == "no", 0, 1)
#drop the original level.up variable 
M1 <- M1 [-7]
```

#Visualization 
```{r}
#open a whole bunch of packages
library(ggplot2)
library(dplyr)
library(tidyr)
#Change data to long form so that 
data.long <-gather (M1,'variable','value',2:7)
M2 <- data.long [-1]
M2$value <-as.numeric(M2$value)
```

```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
ggplot(M2, aes (value,..count..)) + geom_histogram(alpha = 0.5, fill="green", col="blue") + facet_wrap (~variable, scales = 'free')

#Then visualize the relationships between variables
#dropping ID
M3 <- M1 [-1]
library(corrplot)
COR <- cor(M3)
corrplot (COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Try to capture an intution about the data and the relationships
#It seems to me that level up correlates to av.assignement, messages and post.test.socore
# On the other hand, messages are strongly correlated with post.test.score and also the average assignment score so I am not sure what is the third variable here.

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
library(party)

#Plot and generate a CP table for your tree 
c.tree <- rpart(M1$level.up.numeric~ M1$pre.test.score + M1$post.test.score + M1$av.assignment.score, method="class", data=M1) #Notice the standard R notion for a formula X ~ Y

#Look at the error of this tree
printcp(c.tree)
#Root node error: 400/1000 = 0.4
#n= 1000 

#    CP nsplit rel error xerror     xstd
#1 0.93      0      1.00   1.00 0.038730
#2 0.07      1      0.07   0.07 0.013042
#3 0.01      2      0.00   0.00 0.000000
#Plot the tree
post(c.tree, file = "tree.ps", title = "Level Up Graph where 1=Level and 0=Not able to level up")

#Generate a probability value that represents the probability that a student levels up based your classification tree 

M1$pred <- predict(c.tree, type = "prob")[,2]
#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
#install.packages("ROCR")
#install.packages("gplots")
library(ROCR)

#Plot the curve
pred.detail <- prediction(M1$pred, M1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance (pred.detail,"auc"), "y.values" ))
# [1] 1

#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

c.tree2 <- rpart(M1$level.up.numeric~ M1$pre.test.score + M1$forum.posts, method="class", data=M1) 
printcp(c.tree2)
#Root node error: 400/1000 = 0.4
#n= 1000 

#     CP nsplit rel error xerror     xstd
#1 0.3925      0    1.0000 1.0000 0.038730
#2 0.0300      1    0.6075 0.6075 0.033907
#3 0.0200      2    0.5775 0.6100 0.033954
#4 0.0150      3    0.5575 0.5975 0.033716
#5 0.0100      4    0.5425 0.5825 0.033421
#Plot the tree
post(c.tree2, file = "tree2.ps", title = "Level Up Graph where 1=Level and 0=Not able to level up")

#Generate a probability value that represents the probability that a student levels up based your classification tree 

M1$pred2 <- predict(c.tree2, type = "prob")[,2]
#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
#install.packages("ROCR")
#install.packages("gplots")
library(ROCR)

#Plot the curve
pred.detail2 <- prediction(M1$pred2, M1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance (pred.detail2,"auc"), "y.values" ))
# [1] 0.8180375

# Model 1 is so much better, I mean... just look at the graph!!!

```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

M1$threshold.pred1 <- ifelse(M1$pred> 0.9, 1, 0) #very high threadhold
table1 <- table(M1$threshold.pred1, M1$level.up.numeric)
#      0   1
#  0 600   0
#  1   0 400
#Now generate three diagnostics:

D1$accuracy.model1 <- (600+400)/1000
# 100%
D1$precision.model1 <- 400/(400+0)
# 100%
D1$recall.model1 <- 400/(400+0)
# 100%

#Finally, calculate Kappa for your model according to:

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)
#[1] 1.153846


#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

M1$threshold.pred2 <- ifelse(M1$pred> 0.2, 1, 0) #very low threadhold
table2 <- table(M1$threshold.pred2, M1$level.up.numeric)
#      0   1
#  0 600   0
#  1   0 400
#Now generate three diagnostics:

D1$accuracy.model1 <- (600+400)/1000
# 100%
D1$precision.model1 <- 400/(400+0)
# 100%
D1$recall.model1 <- 400/(400+0)
# 100%

#Finally, calculate Kappa for your model according to:

#Convert to matrix
matrix2 <- as.matrix(table2)

#Calculate kappa
kappa(matrix2, exact = TRUE)/kappa(matrix2)
#[1] 1.153846. It's not that much different!!!!
####################################################
#Let's try a different (worse) model?

M1$threshold.pred3 <- ifelse(M1$pred2> 0.9, 1, 0) #very high threadhold
table3 <- table(M1$threshold.pred3, M1$level.up.numeric)
#Finally, calculate Kappa for your model according to:
#Convert to matrix
matrix3 <- as.matrix(table3)
#Calculate kappa
kappa(matrix3, exact = TRUE)/kappa(matrix3)
#[1] 1

### Then....


M1$threshold.pred4 <- ifelse(M1$pred2> 0.2, 1, 0) #very low threadhold
table4 <- table(M1$threshold.pred4, M1$level.up.numeric)
#Finally, calculate Kappa for your model according to:
#Convert to matrix
matrix4 <- as.matrix(table4)
#Calculate kappa
kappa(matrix4, exact = TRUE)/kappa(matrix4)
#[1] 0.9863223

### Finally, we can see some differences here! It's because the first model was too good (overfit?), to show the difference in kappa here model 2 is better.
```


