---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
D1<-read.csv("~/Desktop/HUDK 4050/Assignment#7_Jing/online.data.csv")
```

#Visualization 
```{r}
#data wrangling
library(tidyr)

#remove the id column
D1<-D1[,2:7]

#make the categorical variable numeric
D1$level.up<-ifelse(D1$level.up=="no",0,1)

#since ggplot only deals with data that is in long format
D2<-tidyr::gather(D1,variable,value, 1:6)

#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)

library(ggplot2)

#x-axis in the plot
t <- ggplot(D2,aes(x=value))

#y-axis in the plot
hist= t + geom_histogram(data = D2[D2$variable == "av.assignment.score",], binwidth=0.01) + geom_histogram(data = D2[D2$variable == "forum.posts",], binwidth=1) + geom_histogram(data = D2[D2$variable == "level.up",], binwidth=1) + geom_histogram(data = D2[D2$variable == "messages",], binwidth=1) + geom_histogram(data = D2[D2$variable == "post.test.score",], binwidth=0.01) + geom_histogram(data = D2[D2$variable == "pre.test.score",], binwidth=0.01)+ facet_wrap(~ variable,scales = "free")
hist
#code from Kuang in Github

#hist1 = ggplot(D2,aes(x=value)) + geom_histogram(binwidth = 0.5) + facet_wrap(~variable, scales = "free")
#alternative code from Jon
##why hist1 is different from mine(hist)?

# Most displays are roughly rectangular, so if you have a categorical variable with many levels, it doesn't make sense to try and display them all in one row (or one column). To solve this dilemma, facet_wrap wraps a 1d sequence of panels into 2d, making best use of screen real estate.

#Then visualize the relationships between variables
cor<-cor(D1)
library(corrplot)
corrplot(cor, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.01, insig = "blank")
#code is from assignment#4 pca

#Try to capture an intution about the data and the relationships
#variable post.test.score is highly correlated with variables: messages, av.assignment.score and level.up
#av.assignment.score is highly correlated with level.up and messages
#messages and level.up is correlated 

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
library(party)
c.tree1 <- rpart(level.up ~ av.assignment.score + post.test.score + messages, method="class", data=D1, control=rpart.control(minsplit=1, minbucket=1,cp=0.0001 ))

#Look at the error of this tree; examine the results
printcp(c.tree1)

#Plot and generate a CP table for your tree 
post(c.tree1, file = "tree1.ps", title = "Student Level up")    

#CP nsplit rel error xerror     xstd
#1 0.93      0      1.00   1.00 0.038730
#2 0.07      1      0.07   0.07 0.013042
#3 0.01      2      0.00   0.00 0.000000

#Generate a probability value that represents the probability that a student levels up based your classification tree 

#Based on the classification tree, there are 400 out of 1000 students level up, so the probability is 40%

D1$pred1 <- predict(c.tree1, type = "prob")[,2]
#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail <- prediction(D1$pred1, D1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values")) ###?? what is Pred2???
#Unlist liberates the AUC value from the "performance" object created by ROCR
#[1] 1

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
```

```{r}
#Create model 2 with different variables
c.tree2 <- rpart(level.up ~ pre.test.score + post.test.score + messages, method="class", data=D1, control=rpart.control(minsplit=1, minbucket=1,cp=0.0001 ))

#Look at the error of this tree; examine the results
printcp(c.tree2)

#Plot and generate a CP table for your tree 
post(c.tree2, file = "tree2.ps", title = "Student Level up")  

c.tree2 <- rpart(level.up ~ pre.test.score + forum.posts + messages + av.assignment.score, method="class", data=D1, control=rpart.control(minsplit=1, minbucket=1,cp=0.00500000))

post(c.tree2, file = "tree2.ps", title = "Student Level up")  

D1$pred2 <- predict(c.tree2, type = "prob")[,2]

#Plot the curve
pred.detail2<- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail2,"auc"), "y.values"))
#[1] 0.99525

```
#Thresholds
```{r}
#Look at the ROC plot for your first model.Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

D1$threshold.pred1 <- ifelse(D1$pred1 > 0.8, 1, 0)

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#     0   1
#  0 600   0
#  1   0 400

#Now generate three diagnostics:

D1$accuracy.model1 <- (600+400)/1000
#100%

D1$precision.model1 <- 400/(400 + 0)
#100%

D1$recall.model1 <- 400/(400 + 0)
#100%

#Finally, calculate Kappa for your model according to:

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)
#[1] 1.153846 #what does this mean though? above 1?

###Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

#one accuracy, one precision and one recall for each threshold you test. You should test two thresholds.

###Test another threshold
D1$threshold.pred2 <- ifelse(D1$pred2 > 0.31, 1, 0) 
#predict, not threshold though

table2 <- table(D1$level.up, D1$threshold.pred2)

#      0   1
#  0 595   5
#  1  10 390

#Now generate three diagnostics:

D1$accuracy.model2 <- (595 + 390)/1000
#98.5%

D1$precision.model2 <- 390/(390 + 10)
#97.5%

D1$recall.model2 <- 595/(595 + 10)
#98.35%

#Convert to matrix
matrix2 <- as.matrix(table2)

#Calculate kappa
kappa(matrix2, exact = TRUE)/kappa(matrix2)

#[1] 1.141137 ##why? >1 again?

##To summairize, the model 1 has better prediction, with 100% accuracy, precision and recall, but why the kappa value is above 1???
```


