---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
onlinedata <- read.csv("online.data.csv", header = TRUE, sep = ",")
```

#Visualization 
```{r}
#FOR LATER: try scaling before gather (but after changing the binary yes/no)

#change layout from wide to long (first change yes/no to 1/0):
onlinedata$level.up <- ifelse(onlinedata$level.up=="no", 0,1)
OD <- tidyr::gather(onlinedata, "variable", "value", 2:7)
OD1 <- dplyr::select(OD, -1)
OD1$value <- as.numeric(OD1$value)

library(ggplot2)

#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
p <- ggplot(OD1, aes(value)) + geom_histogram()
p + facet_grid(. ~ variable, scales = "free")
#fixed by charles, work from this. 
#something to keep in mind: `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.

# Error: StatBin requires a continuous x variable the x variable is discrete. Perhaps you want stat="count"?

# GitHub says try geom_bar instead (and that it was the same as histogram, but histogram is for continuous and bar is for discrete).
p <- ggplot(OD1, aes(variable)) + geom_bar()
p + facet_grid(. ~ variable)
#pretty sure I should be using histogram. trying again. 
#this is not the plot I'm looking for: one bar in each that shows count, which is 1000 for each. so the plot shows nothing except that there are 1000 instances of each variable. 

p <- ggplot(OD1, aes(variable, value)) + geom_point()
p + facet_grid(. ~ variable)

#not quite what we're looking for, I don't think. maybe need to scale numbers to use histogram??
OD1$value <- scale(OD1$value, center = TRUE, scale = TRUE)
p <- ggplot(OD1, aes(value)) + geom_histogram()
p + facet_grid(. ~ variable)
#still no. 

#Then visualize the relationships between variables

#Try to capture an intution about the data and the relationships

#messages, post and pre test scores seem to look visually similar in terms of grouping/spread of the plots

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

c.tree <- rpart(level.up ~ post.test.score + pre.test.score + messages, method="class", data=onlinedata) 

#Plot and generate a CP table for your tree 
printcp(c.tree)
#Classification tree:
#rpart(formula = level.up ~ post.test.score + pre.test.score + 
#    messages, data = onlinedata, method = "class")
#
#Variables actually used in tree construction:
#[1] post.test.score
#
#Root node error: 400/1000 = 0.4
#
#n= 1000 
#
#      CP nsplit rel error xerror     xstd
#1 0.6025      0    1.0000 1.0000 0.038730
#2 0.0100      1    0.3975 0.4425 0.030174
plot(c.tree)
#nope

post(c.tree, file = "trying.ps", title = "Indicators: 1 - Something, 2 - Something Else, 3 - Another Thing")

#not what I think it should look like, but definitely something. "Trying"

#Generate a probability value that represents the probability that a student levels up based your classification tree 

onlinedata$pred <- predict(c.tree, type = "prob")[,2]

#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)
#surprisingly tricky. 

#Plot the curve
pred.detail <- prediction(onlinedata$pred, onlinedata$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))
#[1] 0.8545833

#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

c.tree2 <- rpart(level.up ~ messages + forum.posts, method="class", data=onlinedata) 
printcp(c.tree2)
#Classification tree:
#rpart(formula = level.up ~ messages + forum.posts, data = onlinedata, 
#    method = "class")
#
#Variables actually used in tree construction:
#[1] messages
#
#Root node error: 400/1000 = 0.4
#
#n= 1000 
#
#      CP nsplit rel error xerror     xstd
#1 0.5425      0    1.0000 1.0000 0.038730
#2 0.0100      1    0.4575 0.4625 0.030698

post(c.tree2, file = "trying2.ps", title = "Indicators: 1 - Something, 2 - Something Else, 3 - Another Thing")
#go back and figure out why these don't look right later
onlinedata$pred2 <- predict(c.tree2, type = "prob")[,2]

pred.detail2 <- prediction(onlinedata$pred2, onlinedata$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

unlist(slot(performance(pred.detail2,"auc"), "y.values"))
#[1] 0.8379167

#first prediction model looks a little better, but not by a whole lot. Also, only used two variables in second prediction since the level.up is what we're measuring it against...right? First model has a better true/false graph and a slightly better AUC


```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

threshold.pred1 <- 
onlinedata$threshold.pred1 <- ifelse(onlinedata$pred>.62, 1, 0)

odpred <- table(onlinedata$level.up, onlinedata$threshold.pred1) 
odpred

#      0   1
#  0 472 128
#  1  31 369

#Now generate three diagnostics:

onlinedata$accuracy.model1 <- (472+369)/(1000)
# 0.841

onlinedata$precision.model1 <- (369)/(369+31)
# 0.9225
onlinedata$precision2.model1 <- (472)/(472+128)
# 0.7866667

onlinedata$recall.model1 <- (369)/(369+128)
# 0.7424547

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(onlinedata$level.up, onlinedata$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)
# [1] 0.9930247

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

onlinedata$threshold.pred2 <- ifelse(onlinedata$pred2>.3, 1, 0)
odpred2 <- table(onlinedata$level.up, onlinedata$threshold.pred2) 
odpred2
#      0   1
#  0 440 160
#  1  23 377
onlinedata$accuracy.model2 <- (440+377)/1000 
# 0.817
onlinedata$precision.model2 <- (377)/(377+23)
# 0.9425
#I couldn't quite remember what the particulars were for the numbers, so I did both and then realized that wasn't right. In any case, it was good practice. 
onlinedata$precision2.model2 <- (440)/(440+160)
# 0.7333333
onlinedata$recall.model2 <- (377)/(377+160)
# 0.7020484
table2 <- table(onlinedata$level.up, onlinedata$threshold.pred2)
matrix2 <- as.matrix(table2)
kappa(matrix2, exact = TRUE)/kappa(matrix2)
# [1] 0.9061023

#first prediction threshold of .62 seems like a better threshold predictor than .3 when comparing kappas/accuracy/etc. 
```


