---
title: "Assignment 7"
author: "Abdul Abad"
date: "12/18/2018"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
id <- seq(1,1000,1)
D1 <- read.csv("online.data.csv")
D2 <- D1
post.test1 <- rnorm(500, 0.4,0.07)
post.test1 <- ifelse(post.test1 < 0, 0, post.test1)
post.test1 <- ifelse(post.test1 > 1, 1, post.test1)
post.test2 <- rnorm(500, 0.6,0.07)
post.test2 <- ifelse(post.test2 < 0, 0, post.test2)
post.test2 <- ifelse(post.test2 > 1, 1, post.test2)
#post.test3 <- rnorm(300, 0.75,0.07)
#post.test3 <- ifelse(post.test3 > 1, 1, post.test3)
D1$post.test.score <- round(c(post.test1,post.test2),2)
D1$pre.test.score <- round(c(post.test1,post.test2), 2)
D1$messages <- round(sample(c(150:200),1000, replace = TRUE)*D1$post.test.score)
D1$forum.posts <- round(sample(c(0:40),1000, replace = TRUE)*(D1$post.test.score) + 2)
D1$av.assignment.score <- round(D1$post.test.score*rnorm(1000, 0.5,0.1),2)
D1$level.up <- ifelse(D1$post.test.score > 0.45 & D1$av.assignment.score > 0.25, "yes", "no")

```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(ggplot2)
library(dplyr)
library(tidyr)
#HINT: look up "facet" in the ggplot documentation
postscore <- ggplot(D1) + geom_histogram(aes(post.test.score)) + facet_grid(. ~ level.up)
  
prescore <- ggplot(D1) + geom_histogram(aes(pre.test.score)) + facet_grid(. ~ level.up)
messages <- ggplot(D1) + geom_histogram(aes(messages)) + facet_grid(. ~ level.up)
forumposts <- ggplot(D1) + geom_histogram(aes(forum.posts)) + facet_grid(. ~ level.up)
assignmentscore <- ggplot(D1) + geom_histogram(aes(av.assignment.score)) + facet_grid(. ~ level.up)
grid.arrange(postscore, prescore, messages, forumposts, assignmentscore)
D2 <- select(D1, 1:7)
#Convert yes/no to 1/0 to avoid mixing variable types
D2$level.up <- ifelse(D2$level.up == "yes", 1,0)
D3 <- gather(D2, "measure", "score", 2:7)
p1 <- ggplot(D3, aes(score)) + facet_wrap(~measure, scales = "free")
p1 + geom_histogram(stat = "count")

#Then visualize the relationships between variables
pairs(D2)
D2[7] <- ifelse(D2[7] == "yes", 1, 0)
```
COR1 <- cor(D2[2:7])
corrplot(COR1, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank", main = "Online Data")

#Regarding intution, the variables with the least overlap are messages, post test score, average assignment score, which suggests that they will be the best classifiers to figure out when a student level up. Similarly, the correlogram shows how messages, post test score, and av assignment score have the highest correlation with level up.

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

library(rpart)
c.tree1 <- rpart(level.up ~ forum.posts + pre.test.score, method = "class", data = D1, control=rpart.control(minsplit=1, minbucket=1, cp=0.001))
printcp(c.tree1)
plot(c.tree1)
text(c.tree1)


#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred <- predict(rp, type = "prob")[,2]

#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail <- prediction(D1$pred, D1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(Pred2,"auc"), "y.values"))

#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

c.tree2 <- rpart(level.up ~ scale(messages) + scale(av.assignment.score), method="class", data=D1)
printcp(c.tree2)
post(c.tree2, file = "tree2.ps", title = "Level Up 2")
D1$predict2 <- predict(c.tree2, D1, type = "class")
table(D1$level.up, D1$predict2)
D1$pred2 <- predict(c.tree2, type = "prob")[,2]
pred.detail2 <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)
unlist(slot(performance(pred.detail2,"auc"), "y.values"))

#Analysis: 
#Model 1 AUC= 0.8825125
#Model 2 AUC= 0.99525 

#The second model is better because the area under the curve is closer to 1 i.e. it maximizes the true postive rate while minimizing the false positive rate.


```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

D1$threshold.pred1 <- ifelse(D1$pred1 > 0.6124,"yes", "no")

#Now generate three diagnostics:

D1$threshold.pred1 <- ifelse(D1$pred >= 0.8, "yes", "no")
D1$threshold.pred2 <- ifelse(D1$pred >= 0.95, "yes", "no")
D1$threshold.pred3 <- ifelse(D1$pred >= 0.25, "yes", "no")
#Now generate three diagnostics:
accuracy.model1 <- mean(ifelse(D1$level.up == D1$threshold.pred1, 1, 0))
D1$truepos.model1 <- ifelse(D1$level.up == "yes" & D1$threshold.pred1 == "yes", 1, 0)
D1$falsepos.model1 <- ifelse(D1$level.up == "no" & D1$threshold.pred1 == "yes", 1,0)
D1$falseneg.model1 <- ifelse(D1$level.up == "yes" & D1$threshold.pred1 == "no", 1,0)
precision.model1 <- sum(D1$truepos.model1)/(sum(D1$truepos.model1) + sum(D1$falsepos.model1))
recall.model1 <- sum(D1$truepos.model1)/(sum(D1$truepos.model1) + sum(D1$falseneg.model1))
#Finally, calculate Kappa for your model according to:
#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)
table1
#Calculate kappa manually
po <- (586+247)/(586+247+14+153)
pe <- ((586 + 14)/(586+247+14+153))*((586 + 153)/(586+247+14+153)) + ((14 + 247)/(586+247+14+153))*((153+247)/(586+247+14+153))
kappa <- (po - pe)/(1 - pe)
#Calculate OOB
library(psych) #You could also use the "irr" or "vcd" library versions
cohen.kappa(table1)
#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
#Alternate kappa value
library(irr)
kappa2(D1[,c(7,9)], "unweighted")
kappa2(D1[,c(7,10)], "unweighted")
kappa2(D1[,c(7,11)], "unweighted")
```

