---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
programmer: Li Li
Last modified: Dec 2018
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test.score), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
D1 <- read.csv("online.data.csv", header=TRUE)
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(ggplot2)
  ##histgram of post.test.score by level.up 
  ggplot(D1, aes(x=post.test.score)) + geom_histogram(binwidth=0.01, color="white", fill="grey") + facet_grid(.~level.up)
  ##histogram of pre.test.score by level.up
  ggplot(D1, aes(x=pre.test.score)) + geom_histogram(binwidth=0.01, color="white", fill="grey") + facet_grid(.~level.up)
  #histogram of messages by level.up
  ggplot(D1, aes(x=messages)) + geom_histogram(binwidth=2, color="white", fill="grey") + facet_grid(.~level.up)
  ##histogram of forum.posts by level.up
  ggplot(D1, aes(x=forum.posts)) + geom_histogram(binwidth=1, color="white", fill="grey") + facet_grid(.~level.up)
  ##histogram of av.assignment.score by level.up
  ggplot(D1, aes(x=av.assignment.score)) + geom_histogram(binwidth=0.01, color="white", fill="grey") + facet_grid(.~level.up)

#Then visualize the relationships between variables
  library (corrplot)
##remove id and level variables and scale & center variables
  D2 <- D1
  D2$id <- NULL
  D2$level.up <- NULL
  D2 <- scale(D2, center = TRUE)
  D3 <- cbind(D2,as.data.frame(D1$level.up))
  ###rename the 6th column and change it to numeric
  colnames(D3)[6] <- "level.up"
  D3$level.up <- as.numeric(D3$level.up)
##create correlation plot
  COR <- cor(D3)
  corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper", tl.cex=0.6, tl.col="black", tl.srt=45, addCoef.col="white",addCoefasPercent=TRUE, sig.level=0.50, insig="blank")
  
#Try to capture an intution about the data and the relationships
##answer:
###1.Based on the histograms, generally speaking, the distribution of those who get leveled up is more right-ward in all but forum.post variables. In other words, those who get leveled up have higher average pre- and post- score, and higher average of average assignment score and greater average number of messages.
###2.Based on the correlation graph, it seems forum.posts is not a key variable because it is not highly correlated with any of other variables. other variables have at least 0.50 unit correlation with each other. The three variables that are highly correlated with outcome variable ("level.up") are "av.assignment.score", "post.test.score" and "messages".
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
##based on my knowledge and previous descriptive analysis, I would choose post-scores, av.assignment score, and messages
  c.tree <- rpart(level.up ~ post.test.score+pre.test.score+av.assignment.score,  method="class", data=D3)

#Plot and generate a CP table for your tree 
  post(c.tree, file = "leveluptree.ps", title = "5th grade online geography tutoring system")
  printcp(c.tree)
  #note: variables actually used in tree construction: av.assignment.score, post.test.score
  
#Generate a probability value that represents the probability that a student levels up based your classification tree 

D3$pred <- predict(c.tree, D3, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
install.packages("ROCR")
library(ROCR)

#Plot the curve
pred.detail <- prediction(D3$pred, D3$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

##note: auc is 1

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

c.tree2<-rpart(level.up~pre.test.score+forum.posts, method="class",data=D3)

printcp(c.tree2)

post(c.tree2, file = "leveluptree2.ps", title = "5th grade online geography tutoring system")

D3$pred2 <- predict(c.tree2, D3, type="prob")[,2]
 
pred.detail2 <- prediction(D3$pred2, D3$level.up) 

plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

unlist(slot(performance(pred.detail2, "auc"), "y.values"))

##note: auc is 0.8180375

##Model 2 is better. Authough the auc of model 2 is smaller than that of model 1, auc of model 2 is still considered good. Also, since model 1 is overfit for the training data, it may not have good performance for the test data.
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.
##Because the first model has 0 false positive rate, I will analyze the model 2 instead.
D3$threshold.pred2 <- ifelse(D3$pred2 > "0.9", 1, 0)
?ifelse
#Now generate three diagnostics:
##accuracy=(411+27+57+10+278)/1000=0.783
##precision=315/(315+132)=0.705
##recall=315/(315+42+43)=0.7875
D3$accuracy.model2 <- ifelse(D3$pred2 > "0.783", 1, 0)
D3$precision.model2 <-ifelse(D3$pred2 > "0.705", 1, 0) 
D3$recall.model2 <- ifelse(D3$pred2 > "0.7875", 1, 0) 

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D3$level.up, D3$threshold.pred2)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)

##Kappa=1

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
D3$threshold2.pred2 <- ifelse(D3$pred2 > 0.8, 1, 0)
table2 <- table(D3$level.up, D3$threshold2.pred2)
matrix2 <- as.matrix(table2)
kappa(matrix2, exact = TRUE)/kappa(matrix2)
#Kappa=1
```


