---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
D1 <- read.csv("~/Desktop/FALL_2016/Assignment7/online.data.csv")
```

#Visualization 
```{r}
#First, removed the id variable since it had no meanings other than identifying participants, changed level.up variable values into numeric varaible (no=0, yes=1). Then created the cleaned up data to use.
D2 <- dplyr::select(D1,2:7)
D2$level.up2 <- ifelse(D2$level.up == "no",0,1)
D3 <- dplyr::select(D2,1:5,7)
#Then in order to use it for creating ggplot, made the data into long data so that R can understand
data_long <- tidyr::gather(D3,"value_name","value",1:6)
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
1.Started by opening packages.
library(dplyr)
library(tidyr)
library(ggplot2)
2. creating different x and y axis for different variables since they mean different things and want different x/y axis for different variables.
<note to myself: after value_name labeling and putting can comma can be the only way for R to distinguish different variables and their values/ one layers must contain all variables should be value_name this might be the reason why we made the data long to create one histogram>
a <- ggplot(data_long, aes(x=value))
a + geom_histogram(data = data_long[data_long$value_name == "post.test.score",], binwidth=0.01) +
geom_histogram(data = data_long[data_long$value_name == "pre.test.score",], binwidth=0.01) +
geom_histogram(data = data_long[data_long$value_name == "messages",], binwidth=10) +
geom_histogram(data = data_long[data_long$value_name == "forum.posts",], binwidth=2) +
geom_histogram(data = data_long[data_long$value_name == "av.assignment.score",], binwidth=0.01) +
geom_histogram(data = data_long[data_long$value_name == "level.up2",], binwidth=0.01) + 
facet_wrap(~value_name, scales = "free")
#Then visualize the relationships between variables
library(corrplot)
COR <- cor(D3)
corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, 
+ addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")
#Try to capture an intution about the data and the relationships
Starting by the biggest number relationship, messages and post test scores showed the highest relationship. This showes that more interactions students have with other students in the class, students got the better grade. (Not to be suspicious, but I was wondering if students shared the answeres through message since this is online course; just my second thought). Also average assignment score and level up status was positively correlated. I coded level up yes as 1, so this relationship is understandable and makes sense; students who performs well might need advance education in geography. I was just interested in why forum post and message dose not have a high correlation since both are the outlet for students to communicate...
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

1. open relative package
library(rpart)
library(party)

2.c.tree1 <- rpart(level.up2 ~ messages + post.test.score + pre.test.score, method="class", data=D3, control=rpart.control(minsplit=1,minbucket=1,cp=0.001))

#Plot and generate a CP table for your tree 
printcp(c.tree1)
Classification tree:
rpart(formula = level.up2 ~ messages + post.test.score + pre.test.score, 
    data = D3, method = "class", control = rpart.control(minsplit = 1, 
        minbucket = 1, cp = 0.001))

Variables actually used in tree construction:
[1] messages        post.test.score pre.test.score 

Root node error: 400/1000 = 0.4

n= 1000 

          CP nsplit rel error xerror     xstd
1  0.6025000      0    1.0000 1.0000 0.038730
2  0.0062500      1    0.3975 0.4400 0.030106
3  0.0050000      4    0.3775 0.4425 0.030174
4  0.0041667      5    0.3725 0.4700 0.030889
5  0.0037500      8    0.3600 0.4825 0.031200
6  0.0033333     14    0.3375 0.4875 0.031322
7  0.0025000     22    0.3025 0.4875 0.031322
8  0.0018750     77    0.1575 0.4825 0.031200
9  0.0016667     81    0.1500 0.5000 0.031623
10 0.0012500    112    0.0975 0.5000 0.031623
11 0.0010714    162    0.0350 0.5200 0.032087
12 0.0010000    171    0.0250 0.5200 0.032087

3.code for tree ploting
post(c.tree1, file = "tree.ps", title = "level_up_decision")

#Generate a probability value that represents the probability that a student levels up based your classification tree 
D3$pred1 <- predict(c.tree1, type = "prob")[,2]

#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred1 <- prediction(D3$pred1, D3$level.up2) 
plot(performance(pred1, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred1,"auc"), "y.values")) #Unlist liberates the AUC value from the "performance" object created by ROCR

[1] 0.998925

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

As I mentioned previously I was interested in the difference between message and the forum post, so this time I used forum posts, average assignment score, and pre test score variable. 

There were no significant difference in the values only 0.0016 differences, so either one is fine.

c.tree2 <- rpart(level.up2 ~ forum.posts + av.assignment.score + pre.test.score, method="class", data=D3, control=rpart.control(minsplit=1,minbucket=1,cp=0.001))
printcp(c.tree2)

Classification tree:
rpart(formula = level.up2 ~ forum.posts + av.assignment.score + 
    pre.test.score, data = D3, method = "class", control = rpart.control(minsplit = 1, 
    minbucket = 1, cp = 0.001))

Variables actually used in tree construction:
[1] av.assignment.score forum.posts         pre.test.score     

Root node error: 400/1000 = 0.4

n= 1000 

         CP nsplit rel error xerror     xstd
1 0.9300000      0    1.0000 1.0000 0.038730
2 0.0075000      1    0.0700 0.0700 0.013042
3 0.0037500      3    0.0550 0.0600 0.012100
4 0.0025000      7    0.0375 0.0675 0.012814
5 0.0016667     13    0.0225 0.0525 0.011336
6 0.0012500     16    0.0175 0.0525 0.011336
7 0.0010000     18    0.0150 0.0525 0.011336
post(c.tree2, file = "tree.ps2", title = "level_up_decision2")
D3$pred2 <- predict(c.tree2, type = "prob")[,2]
pred2 <- prediction(D3$pred2, D3$level.up2) 
plot(performance(pred2, "tpr", "fpr"))
abline(0, 1, lty = 2)
unlist(slot(performance(pred2,"auc"), "y.values"))
[1] 0.9973688
```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.
I have chosen the value of 0.9 since most of the variables are to the extreme...
D3$threshold.pred1 <- ifelse(D3$pred1 > 0.9, 1, 0)
table(D3$threshold.pred1, D3$level.up2)
   
      0   1
  0 596  22
  1   4 378

#Now generate three diagnostics:

D3$accuracy.model1 <- 
(596+378)/(596+378+22+4) = 0.974                                                                                                                      
D3$precision.model1 <-
(378) / (378+22) = 0.945

D3$recall.model1 <- 
(378) / (378+4) = 0.9895288

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D3$threshold.pred1, D3$level.up2)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)
[1] 1.130513

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

I have chosen the value of 0.95 since most of the variables are to the extreme...

D3$threshold.pred2 <- ifelse(D3$pred1 > 0.95, 1, 0)
table(D3$threshold.pred2, D3$level.up2)
   
    0   1
  0 599  62
  1   1 338

#Now generate three diagnostics:

D3$accuracy.model2 <- 
(599+338)/(599+1+62+338) = 0.937                                                                                                                   
D3$precision.model2 <-
(338) / (338+62) = 0.845

D3$recall.model2 <- 
(338) / (338+1) = 0.9970501

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table2 <- table(D3$threshold.pred2, D3$level.up2)

#Convert to matrix
matrix2 <- as.matrix(table2)

#Calculate kappa
kappa(matrix2, exact = TRUE)/kappa(matrix2)
1.102105
```



