---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
D1<-read.csv("online.data.csv",header=TRUE)
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(ggplot2)
library(tidyr)
D2<-D1
D2$level.up<-ifelse(D2$level.up=="yes",1,0)
D3<-gather(D2,"variable","value",2:7)
ggplot(D3,aes(value))+facet_wrap(~variable,scale="free")+geom_histogram(stat="count")

#Then visualize the relationships between variables
pairs(D2)
## Post test score and average assignment score are positively related to the number of messages sent by each student to other students about the assignments.

#Try to capture an intution about the data and the relationships
## Students who send more messages about the assignment are more likely to obtain more information and knowledge about the assignment, thus getting higher scores.
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
c.tree1<-rpart(as.factor(level.up) ~ forum.posts + pre.test.score + post.test.score, method = "class", data = D1,control=rpart.control(minsplit=1, minbucket=1, cp=0.001))

#Plot and generate a CP table for your tree 
printcp(c.tree1)
plot(c.tree1)
text(c.tree1, cex=0.6)

#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred1 <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
library(ROCR)

#Plot the curve
pred.detail1 <- prediction(D1$pred1, D1$level.up) 
plot(performance(pred.detail1, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail1,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
c.tree2<-rpart(as.factor(level.up) ~ av.assignment.score + messages, method = "class", data = D1,control=rpart.control(minsplit=1, minbucket=1, cp=0.001))

printcp(c.tree2)
plot(c.tree2)
text(c.tree2, cex=0.6)

D1$pred2 <- predict(c.tree2, type = "prob")[,2]

pred.detail2<- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

unlist(slot(performance(pred.detail2,"auc"), "y.values"))


performance(pred.detail2,"acc")
## The first model is better since the AUC value of the first model is a litter higher, which means the probability that the classifier will score are randomly drawn positive sample higher than a randomly drawn negative sample. Larger values of AUC are related to higher estimate-observation agreement.
## The smoother the graph, the more cutoffs the predictions have. I also plotted a 45-degree line, which represents, on average, the performance of a Uniform(0, 1) random variable. The further away from the diagonal line, the better. Overall, I see that the predictions are similar.
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

D1$threshold.pred1 <- ifelse(D1$pred1 >=0.8, 1, 0)

#Now generate three diagnostics:
table <- table(D1$threshold.pred1, D1$level.up)
table
D1$accuracy.model1 <- sum(diag(table))/sum(table)    ## 0.984

D1$precision.model1 <- table[2,2]/sum(table[2,])     ## 0.9752475

D1$recall.model1 <- table[2,2]/sum(table[,2])        ## 0.985

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)  ## 1.133458
```

```{r}
#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
D1$threshold.pred2 <- ifelse(D1$pred1 >= 0.9, "yes", "no")
D1$accuracy.model2 <- mean(ifelse(D1$level.up == D1$threshold.pred2, 1, 0))            ## 0.959
D1$truepos.model2 <- ifelse(D1$level.up == "yes" & D1$threshold.pred2 == "yes", 1, 0)
D1$falsepos.model2 <- ifelse(D1$level.up == "no" & D1$threshold.pred2 == "yes", 1,0)
D1$falseneg.model2 <- ifelse(D1$level.up == "yes" & D1$threshold.pred2 == "no", 1,0)
D1$precision.model2 <- sum(D1$truepos.model2)/(sum(D1$truepos.model2) + sum(D1$falsepos.model2)) ## 0.9838275
D1$recall.model2 <- sum(D1$truepos.model2)/(sum(D1$truepos.model2) + sum(D1$falseneg.model2))   ## 0.9125

table2 <- table(D1$level.up, D1$threshold.pred2)

#Convert to matrix
matrix2 <- as.matrix(table2)

#Calculate kappa
kappa(matrix2, exact = TRUE)/kappa(matrix2)    ## 1.13268


## Results of two thresholds are similar.Kappa of the first threshold is a little higher, which means the extent of agreement among data collectors is called, “interrater reliability”, is higher. The accuracy of the first threshold is also a little higher than that of the second threshold.
```

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.
