---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
M1 <- read.csv("online.data.csv")
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(ggplot2)
library(dplyr)
library(tidyr)
M1$level.up<-ifelse(M1$level.up=="yes",1,0)

#Then visualize the relationships between variables
#We have to using gather fuction in order to use facet_wrap
M2<-gather(M1,"Key","Value",2:7)

ggplot(M2,aes(Value))+geom_histogram()+facet_wrap(~Key,scale="free")



#Try to capture an intuition about the data and the relationships
pairs(M1)
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
#first I want to figure out which variables are more related.
library(rpart)
c.tree1<-rpart(factor(level.up)~.,method="class", data=M1)
printcp(c.tree1)
post(c.tree1, file = "tree1.ps", title = "levels up")

#actually there are only two variables av.assignment.score and post.test.score is accountable for prediction level up. 
# if I choose av.assignment.score+post.test.score, the error rate will comes to zero.
c.tree2<-rpart(factor(level.up)~forum.posts+av.assignment.score+post.test.score,method="class",data=M1)
printcp(c.tree2)
post(c.tree2,file="tree2.ps",title="levels up")
#Plot and generate a CP table for your tree 

#Generate a probability value that represents the probability that a student levels up based your classification tree 

M1$pred <- predict(c.tree2, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
#what does [,2] mean?
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
library(ROCR)

#Plot the curve
pred.detail <- prediction(M1$pred, M1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))+abline(0, 1, lty = 2,)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
c.tree3<-rpart(factor(level.up)~forum.posts+pre.test.score+messages,method="class",data=M1)
printcp(c.tree3)
M1$pred1 <- predict(c.tree3, type = "prob")[,2]
pred.detail1<-prediction(M1$pred1, M1$level.up)
plot(performance(pred.detail1, "tpr", "fpr"))+abline(0, 1, lty = 2,)
unlist(slot(performance(pred.detail1,"auc"), "y.values"))
# The auc value goes down from 1 to 0.88, which means the accuracy drops using the second value.
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

#Since the first accuracy rate is 100%. It is not a unusal case, I will use the second model which the accruacy rate is not 100% 

#I will try 0.5 & 0.6
M1$threshold.pred1 <-ifelse(M1$pred1>0.5,1,0)

M1$threshold.pred2<-ifelse(M1$pred1>0.7,1,0)

#Now generate three diagnostics:

accuracy.model1 <-mean(ifelse(M1$threshold.pred1==M1$level.up,1,0))
accuracy.model1
accuracy.model2<-mean(ifelse(M1$threshold.pred2==M1$level.up,1,0))
accuracy.model2
#model1 accuracy is 0.826 while model2 accuracy is 0.772

#I need to figure out true positive,false positive, false positive in order to caculate 
#for threshold 1
M1$truepos1<-ifelse(M1$level.up==1&M1$threshold.pred1==1,1,0)
M1$falsepos1<-ifelse(M1$level.up==0&M1$threshold.pred1==1,1,0)
M1$falseneg1<-ifelse(M1$level.up==1&M1$threshold.pred1==0,1,0)

#for threshold 2
M1$truepos2<-ifelse(M1$level.up==1&M1$threshold.pred2==1,1,0)
M1$falsepos2<-ifelse(M1$level.up==0&M1$threshold.pred2==1,1,0)
M1$falseneg2<-ifelse(M1$level.up==1&M1$threshold.pred2==0,1,0)


M1$precision.model1 <-sum(M1$truepos1)/(sum(M1$truepos1)+sum(M1$falsepos1))
#model1 precision is 0.7306
M1$precision.model2 <-sum(M1$truepos2)/(sum(M1$truepos2)+sum(M1$falsepos2))
#model 2 precision is 0.844
M1$recall.model1 <-sum(M1$truepos1)/(sum(M1$truepos1)+sum(M1$falseneg1))
#model1 recall is 0.895
M1$recall.model2 <-sum(M1$truepos2)/(sum(M1$truepos2)+sum(M1$falseneg2))
#model2 recall is 0.5275
#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(M1$level.up, M1$threshold.pred1)
table1
table2 <- table(M1$level.up, M1$threshold.pred2)
table2
#Convert to matrix
matrix1 <- as.matrix(table1)
matrix2<-as.matrix(table2)
#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)
kappa(matrix2, exact = TRUE)/kappa(matrix2)
# kappa is not possible >1 
#install.packages("psych")
library(psych) 
cohen.kappa(matrix1)
cohen.kappa(matrix2)

#model 1 have the higher kappa. 
#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
#answer shows above？
```

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.
