---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
library(ggplot2)
library(dplyr)
library(corrplot)
library(rpart)
setwd("~/Nino Zhang/study at TC/fall 2016/EDM/assig10/assignment7")

data1 <- read.csv("online.data.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
# change character vars "no", "yes" into 0 and 1
data1$level.up <- sub("^(no).*", "0", data1$level.up)
data1$level.up <- sub("^(yes).*", "1", data1$level.up)
class(data1$level.up)
# change character vars into numeric vars
data1$level.up <- as.numeric(data1$level.up)
class(data1$level.up)
datanoid <- select(data1, -1)
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
hist(data1$post.test.score, xlab = "post test score", main = "post test score")
hist(data1$pre.test.score, xlab = "pret test score", main = "pre test score")
hist(data1$messages, xlab = "messages", main = "messages")
hist(data1$forum.posts, xlab = "forum post", main = "forum posts")
hist(data1$av.assignment.score, xlab = "Average score", main = "avg score")
hist(data1$level.up, xlab = "level up", main = "level up")

#Then visualize the relationships between variables
#Try to capture an intution about the data and the relationships
cor_pairs <- cor(datanoid)
corrplot(cor_pairs)

```
The bigger the circle gets, the higher the correlation. Thus, messages have a high correlation with post test score, forum posts have very low correlation with other factors. 

#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
post_test_score <- datanoid$post.test.score
pre_test_score <- datanoid$pre.test.score
messages <- datanoid$messages
forum_posts <- datanoid$forum.posts
avg_score <- datanoid$av.assignment.score
level_up <- datanoid$level.up
c.tree1 <- rpart(level_up ~ forum_posts + messages + pre_test_score, method="class", control = rpart.control(minsplit = 1, minbucket = 1, cp = 0.001), data=datanoid)
#Plot and generate a CP table for your tree 
printcp(c.tree1)
plot(c.tree1)
#Generate a probability value that represents the probability that a student levels up based your classification tree 
datanoid$pred <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
datanoid$pred[1:20]
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail <- prediction(datanoid$pred, datanoid$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
```
```{r}
#repeat the previous process
c.tree2 <- rpart(level_up ~ post_test_score + avg_score, method="class", control = rpart.control(minsplit = 1, minbucket = 1, cp = 0.001), data=datanoid)
#Plot and generate a CP table for your tree 
printcp(c.tree2)
plot(c.tree2)
#Generate a probability value that represents the probability that a student levels up based your classification tree 
datanoid$pred2 <- predict(c.tree2, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
datanoid$pred2[1:20]
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

#Plot the curve
pred.detail2 <- prediction(datanoid$pred2, datanoid$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))

```
Since both AUCs of the two models are 0.9961771, they have the same performance.

#Thresholds
```{r, warning=FALSE, echo=TRUE, include=FALSE}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

#datanoid$threshold.pred1 <- ifelse(datanoid$pred> 0.7, 1, 0)

#Now generate three diagnostics:
#table1 <- table(datanoid$threshold.pred1, level_up)
#table1
#datanoid$accuracy.model1 <- (584+397)/1000

#datanoid$precision.model1 <- 584/(584+16)

#datanoid$recall.model1 <- 584/584+3

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
#table1 <- table(datanoid$level.up, datanoid$threshold.pred1)

#Convert to matrix
#matrix1 <- as.matrix(table1)

#Calculate kappa
#kappa(matrix1, exact = TRUE)/kappa(matrix1)

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
```
Kappa should fall between 0 and 1, maybe it's because of the threshold
```{r, warning=FALSE, echo=TRUE, include=FALSE}
#datanoid$threshold.pred1_2 <- ifelse(datanoid$pred> 0.001, 1, 0)

#Now generate three diagnostics:
#table2 <- table(datanoid$threshold.pred1_2, level_up)
#table2
#datanoid$accuracy.model2 <- (579+398)/1000

#datanoid$precision.model2 <- 579/(579+21)

#datanoid$recall.model2 <- 579/579+2

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
#table2 <- table(datanoid$level.up, datanoid$threshold.pred1_2)

#Convert to matrix
#matrix2 <- as.matrix(table2)

#Calculate kappa
#kappa(matrix2, exact = TRUE)/kappa(matrix2)
```
Kappa is still bigger than 1!!!
```{r, warning=FALSE, echo=TRUE, include=FALSE}
#datanoid$threshold.pred1_3 <- ifelse(datanoid$pred2> 0.00001, 1, 0)

#Now generate three diagnostics:
#table3 <- table(datanoid$threshold.pred1_3, level_up)
#table3
#datanoid$accuracy.model3 <- (600)/1000

#datanoid$precision.model3 <- 600/(600+1)

#datanoid$recall.model3 <- 600/600

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
#table3 <- table(datanoid$level.up, datanoid$threshold.pred1_3)

#Convert to matrix
#matrix3 <- as.matrix(table3)

#Calculate kappa
#kappa(matrix3, exact = TRUE)/kappa(matrix3)
```
Kappa is still bigger than 1...?






##Tried the kappa2 function from irr package to compute kappa value.




Take two, with the irr package
```{r, warning=FALSE, echo=TRUE, include=TRUE}
library(irr)
kappa2(datanoid[,c(6,7)], "unweighted")
```

Here, kappa = 0.638, which is between 0 and 1. 
