---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
D1 <- read.csv('/Users/Gemy/assignment7/online.data.csv')
###Recode D1 level up to binary
D1$level.up <- ifelse(D1$level.up == 'yes', 1, 0)

```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)

library(ggplot2)


### Change to data long format

library(tidyr)

data_long <- gather(D1,'vars', 'value', 2:7)

D2 <- data_long[-1]
D2$value <- as.numeric(D2$value)

```

# Plotting the histograms
```{r}
ggplot(D2,aes(value, ..density..)) + geom_histogram(binwidth = .1) + facet_wrap(~vars, scales = 'free')
```

#Then visualize the relationships between variables:
```{r}

### First take a look at correlation between variables:

cor(D1)

pairs(D2)

### Take a look at high correlations in details with bigger plots
## Messages on Pre-Test Scores 

qplot(x = messages, y = pre.test.score, data = D1, main = "Message on Pre-Test Scores", margins = TRUE)
## Messages on Post-Test Scores

qplot(x = messages, y = post.test.score, data = D1, main = "Message on Post-Test Scores", margins = TRUE)

## Messages on Average Assignment Score

qplot(x = messages, y = av.assignment.score, data = D1, main = "Messages on Average Assignment Score", margins = TRUE)

##Average Assignment Score on Post Test Scores:
qplot(x = av.assignment.score, y = post.test.score, data = D1, main = "Average Assignment Score on Post Test Scores", margins = TRUE)


```
#Try to capture an intution about the data and the relationships

# We notice high correlaion especially between number of messages and pre and post test scores, also number of messages and average assignment scores.
#Average Assignment Score also seems to have high correlation with Post Test scores

#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

#Use rpart package to generate a classification tree:
library(rpart)
tree1 <- rpart(level.up ~ messages + forum.posts + pre.test.score, method ='class', control = rpart.control(minsplit =1, minbucket =1, cp =0.001), data = D1)


#Plot and generate a CP table for your tree 

printcp(tree1)

#Plot your tree
post(tree1, file = "tree1.ps", title = "Students Level Up or NOT")

#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred <- predict(tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail <- prediction(D1$pred, D1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
tree2 <- rpart(level.up ~ av.assignment.score + post.test.score, method ='class', control = rpart.control(minsplit =1, minbucket =1, cp =0.001), data = D1)
printcp(tree2)
post(tree2, file ='tree2.ps', title ="Students Level Up or Not Version 2")
D1$pred2 <- predict(tree2, type ='prob')[,2]

## Now generate ROC curve for the new model:
#Plot the curve
pred.detail2 <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail2,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR


##The second a  better model because the auc value is exactly 1, which indicates a better classifier/ better model. Regardless the first model is also very good with auc = 0.996

```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

## According to the first ROC plot for my first model, it shows the probability threshold that balances captures the most correct prediction against fall positives are 0.1

D1$threshold.pred1 <- ifelse(D1$pred > 0.1, 1, 0)



#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)

## Kappa for matrix 1 = 1.103


```
## Generate 3 Diagnostics: 
```{r}
table1
matrix1

```

```{r}

#Now generate three diagnostics for the first model: 

D1$accuracy.model1 <- (571+398)/(571+29+2+398)
#accuracy = 0.969 = 97% accuracy 
D1$precision.model1 <- 571/(571+2)
#precision =0/9966 = 99% = percentage of the prediction that are correct
D1$recall.model1 <- 571/(571+29)
#recall = 0.9517 = 95% of the positives/correct instances come from the model 


```

```{r}
#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

D1$threshold.pred2 <- ifelse(D1$pred > 0.9, 1, 0)

table2 <- table(D1$level.up, D1$threshold.pred2)
              
#Convert to matrix
matrix2 <- as.matrix(table2)

#Calculate kappa
kappa(matrix2, exact = TRUE)/kappa(matrix2)

## Kappa for matrix 2 = 1.128

```

#Now generate three diagnostics for the second model: 

```{r}


D1$accuracy.model2 <- (594+346)/(594+6+54+346)
## accuracy = 0.94 = 94% accuracy 
D1$precision.model2<- 594/(594+54)
## precision =0.917 = 92% precision
D1$recall.model2<- 594/(594+6)
## recall = 0.99 = 99% corrects come from this model 
```

  