---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
DF1 <- read.csv("online.data.csv", header = TRUE)
View(DF1)

library(reshape2)
library(ggplot2)
library(dplyr)
library(tidyr)
library(corrplot)
library(rpart)

```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
?Facet
#Using facet, the data is split up by the level.up variable on the horizontal direction.
#Then visualize the relationships between variables
#Try to capture an intution about the data and the relationships

HP1 <- ggplot(DF1, aes(x=post.test.score)) + geom_histogram(binwidth=0.1,colour="white") + facet_grid(~level.up)
plot(HP1)
#You would expect to see students that scored above averge leveling up (as reflected in the socres for those that did level up) but you do not see this reflected here.

HP2 <- ggplot(DF1, aes(x=av.assignment.score)) + geom_histogram(binwidth=0.05,colour="white") + facet_grid(~level.up)
plot(HP2)
#Students that scored higher are more likely to level up, which better aligns with the rletaionship one would expect to see.

HP3 <- ggplot(DF1, aes(x=pre.test.score)) + geom_histogram(binwidth=0.1,colour="white") + facet_grid(~level.up)
plot(HP3)

HP4 <- ggplot(DF1, aes(x=messages)) + geom_histogram(binwidth=10,colour="white") + facet_grid(~level.up)
plot(HP4)

HP5 <- ggplot(DF1, aes(x=forum.posts)) + geom_histogram(binwidth=2,colour="white") + facet_grid(~level.up)
plot(HP5)

# The amount of messages and posts is higher for those that did not level up. If we assume that this level of engagemnet is motivated by the need for added or targetted support then it is possible that those that need more help may not be able to level up within the observed period.
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

library(rpart)
c.tree1 <- rpart(level.up ~ av.assignment.score+ post.test.score, control = rpart.control(minsplit=15 ,cp=0.001), method ="class", data =DF1)

#Plot and generate a CP table for your tree 
printcp(c.tree1)
plot(c.tree1)
summary(c.tree1)
post(c.tree1, file = "tree1.ps", title = "STUDENT LEVEL PREDICTION")
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
#Generate a probability value that represents the probability that a student levels up based your classification tree 

library(rpart)
c.tree1 <- rpart(level.up ~ av.assignment.score+ post.test.score, control = rpart.control(minsplit=15 ,cp=0.001), method ="class", data =DF1)
DF1$pred <- predict(c.tree1, type = "prob")[,2]
#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Plot the curve
install.packages("ROCR") 
library(ROCR)

DF1$pred <- predict(c.tree1, type = "prob")[,2]
pred.detail <- prediction(DF1$pred, DF1$level.up)
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
#Unlist liberates the AUC value from the "performance" object created by ROCR
unlist(slot(performance(pred.detail,"auc"), "y.values"))

```

```{r}
#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. 

library(ROCR)
library(rpart)
c.tree2 <- rpart(level.up ~ pre.test.score + forum.posts + messages, control = rpart.control(minsplit = 25, cp=0.001),method="class", data = DF1)

printcp(c.tree2)
DF1$pred <- predict(c.tree2, type = "prob")[,2]
pred.detail2 <- prediction(DF1$pred, DF1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)
unlist(slot(performance(pred.detail2,"auc"), "y.values"))
```

```{r}
#Which one do you think was the better model? Why?
unlist(slot(performance(pred.detail,"auc"), "y.values"))
unlist(slot(performance(pred.detail2,"auc"), "y.values"))

#Model 1 is a better model given that the area under the ROC curve is larger in comparison to that of model 2. The ROC curvey for model 1 has an area of 1, which represents a perfect test. 
```

## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

```

```{r}
threshold.pred.yes <- ifelse(DF1$pred>=.50,1,0)
leveluppositive <- ifelse(DF1$level.up=="yes", 1,0)
threshold.pred.no <- 1000 - sum(threshold.pred.yes)
levelupnegative <- 1000 - sum(leveluptrue)

#Now generate three diagnostics:
# To calculate the true positive, you take the predicted yes count and divide that by the total number of students that leveled up in the data set. To generate the true negative, you subtract the predicted yes and total yes from the total count of students (1000) and what remains accounts for the predicted no and the total count of students that did not level up.

true.positive<- sum(threshold.pred.yes)/sum(leveluppositive)
true.negative<- threshold.pred.no/levelupnegative
false.positive<- threshold.pred.no/sum(leveluppositive)
false.negative<- sum(threshold.pred.yes)/levelupnegative

# Accuracy accounts for the number of correct predictions over the total number of predictions.
DF1$accuracy.model1 <- (true.positive+true.negative)/1000
unique(DF1$accuracy.model1)

# Precision accounts for the number of true positives over the summation of true positives and false positives. 
DF1$precision.model1 <- true.positive/(true.positive+false.positive)
unique(DF1$precision.model1)

# Recall accounts for the number of true positives over the summation of true positives and false negatives. 
DF1$recall.model1 <- true.positive/(true.positive+false.negative)
unique(DF1$recall.model1)

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(leveluppositive, threshold.pred.yes)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)
```

```{r}
#Now choose a different threshold value and repeat these diagnostics. 

threshold.pred.yes2 <- ifelse(DF1$pred>=.60,1,0)
leveluptrue2 <- ifelse(DF1$level.up=="yes", 1,0)
threshold.pred.no2 <- 1000 - sum(threshold.pred.yes)
levelupnegative2 <- 1000 - sum(leveluptrue)

true.positive<- sum(threshold.pred.yes2)/sum(leveluptrue2)
true.negative<- threshold.pred.no2/levelupnegative2
false.positive<- threshold.pred.no2/sum(leveluptrue2)
false.negative<- sum(threshold.pred.yes2)/levelupnegative2

# Accuracy accounts for the number of correct predictions over the total number of predictions.
DF1$accuracy.model2 <- (true.positive+true.negative)/1000
unique(DF1$accuracy.model2)

# Precision accounts for the number of true positives over the summation of true positives and false positives. 
DF1$precision.model2 <- true.positive/(true.positive+false.positive)
unique(DF1$precision.model2)

# Recall accounts for the number of true positives over the summation of true positives and false negatives. 
DF1$recall.model2 <- true.positive/(true.positive+false.negative)
unique(DF1$recall.model2)

#Generate the table of comparisons
table2 <- table(leveluptrue2, threshold.pred.yes2)
matrix2 <- as.matrix(table1)

#Calculate kappa
kappa(matrix2, exact = TRUE)/kappa(matrix2)

#What conclusions can you draw about your two thresholds? The first threshhold was set to >=50% and the second to >=60%. Accuracy, precision and Kappa decrease from threshold one to threshold two. Meanwhile, recall remains the same across both thresholds. 

```