---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---
Assignment 7- Euki Sawada

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I
```{r}
install.packages("plotly")
install.packages("ROCR")
install.packages("irr")
```


```{r}
library(ggplot2)
library(dplyr)
library(rpart)
library(party)
library(reshape2)
library(plotly)
library(gridExtra)
library(corrplot)
library(irr)
```

#Upload data
```{r}
D1<-read.csv("online.data.csv")
D2 <- D1
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)

postscore <- ggplot(D1) + geom_histogram(aes(post.test.score)) + facet_grid(. ~ level.up)
  
prescore <- ggplot(D1) + geom_histogram(aes(pre.test.score)) + facet_grid(. ~ level.up)

messages <- ggplot(D1) + geom_histogram(aes(messages)) + facet_grid(. ~ level.up)

forumposts <- ggplot(D1) + geom_histogram(aes(forum.posts)) + facet_grid(. ~ level.up)

assignmentscore <- ggplot(D1) + geom_histogram(aes(av.assignment.score)) + facet_grid(. ~ level.up)

grid.arrange(postscore, prescore, messages, forumposts, assignmentscore)

#Then visualize the relationships between variables
D2[7] <- ifelse(D2[7] == "yes", 1, 0)

COR1 <- cor(D2[2:7])
corrplot(COR1, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank", main = "Online Data")

#Try to capture an intution about the data and the relationships
```
There seems to be the least overlap between level up histograms for messages, post test score, and average assignment score. This indicates these three variables are likely to be the most significant classifiers for determining whether a student levels up or not. The correlogram also reflects this conclusion since messages, post.test.score, and av.assignment.score have the highest correlation with level.up.

#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

c.tree1 <- rpart(level.up ~ scale(forum.posts) + scale(pre.test.score) + scale(messages), method="class", data=D1)

printcp(c.tree1)
post(c.tree1, file = "tree1.ps", title = "Level Up")

#Plot and generate a CP table for your tree 
D1$predict1 <- predict(c.tree1, D1, type = "class")
table(D1$level.up, D1$predict1)

#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred1 <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
library(ROCR)

#Plot the curve
pred.detail1 <- prediction(D1$pred1, D1$level.up) 
plot(performance(pred.detail1, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail1,"auc"), "y.values")) #Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

c.tree2 <- rpart(level.up ~ scale(messages) + scale(av.assignment.score), method="class", data=D1)

printcp(c.tree2)

post(c.tree2, file = "tree2.ps", title = "Level Up 2")

D1$predict2 <- predict(c.tree2, D1, type = "class")

table(D1$level.up, D1$predict2)

D1$pred2 <- predict(c.tree2, type = "prob")[,2]

pred.detail2 <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

unlist(slot(performance(pred.detail2,"auc"), "y.values"))
```
Part II Analysis: 
Model 1 AUC= 0.8825125
Model 2 AUC= 0.99525 

The second model is better because the area under the curve is closer to 1 i.e. it maximizes the true postive rate while minimizing the false positive rate.

## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

D1$threshold.pred1 <- ifelse(D1$pred1 > 0.6124,"yes", "no")

#Now generate three diagnostics:
D1$correctpred1 <- ifelse(D1$threshold.pred1 == D1$level.up, 1, 0)
D1$tp.pred1 <- ifelse(D1$threshold.pred1=="yes"&D1$level.up=="yes",1,0)
D1$fp.pred1 <- ifelse(D1$threshold.pred1=="yes"&D1$level.up=="no",1,0)
D1$fn.pred1 <- ifelse(D1$threshold.pred1=="no"&D1$level.up=="yes",1,0)

accuracy1 <- mean(D1$correctpred1)
precision1 <- sum(D1$tp.pred1)/(sum(D1$tp.pred1)+sum(D1$fp.pred1))
recall1 <- sum(D1$tp.pred1)/(sum(D1$tp.pred1)+sum(D1$fn.pred1))

#Finally, calculate Kappa for your model according to:
kappa2(select(D1, level.up, threshold.pred1))

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

#Threshold 2 (0.8440)
D1$threshold.pred2 <- ifelse(D1$pred1 >= 0.8440,"yes", "no")

D1$correctpred2 <- ifelse(D1$threshold.pred2 == D1$level.up, 1, 0)
D1$tp.pred2 <- ifelse(D1$threshold.pred2=="yes"&D1$level.up=="yes",1,0)
D1$fp.pred2 <- ifelse(D1$threshold.pred2=="yes"&D1$level.up=="no",1,0)
D1$fn.pred2 <- ifelse(D1$threshold.pred2=="no"&D1$level.up=="yes",1,0)

accuracy2 <- mean(D1$correctpred2)
precision2 <- sum(D1$tp.pred2)/(sum(D1$tp.pred2)+sum(D1$fp.pred2))
recall2 <- sum(D1$tp.pred2)/(sum(D1$tp.pred2)+sum(D1$fn.pred2))

kappa2(select(D1, level.up, threshold.pred2))

#Threshold 3 (0.4042)
D1$threshold.pred3 <- ifelse(D1$pred1 >= 0.4042,"yes", "no")

#Now generate three diagnostics:
D1$correctpred3 <- ifelse(D1$threshold.pred3 == D1$level.up, 1, 0)
D1$tp.pred3 <- ifelse(D1$threshold.pred3=="yes"&D1$level.up=="yes",1,0)
D1$fp.pred3 <- ifelse(D1$threshold.pred3=="yes"&D1$level.up=="no",1,0)
D1$fn.pred3 <- ifelse(D1$threshold.pred3=="no"&D1$level.up=="yes",1,0)

accuracy3 <- mean(D1$correctpred3)
precision3 <- sum(D1$tp.pred3)/(sum(D1$tp.pred3)+sum(D1$fp.pred3))
recall3 <- sum(D1$tp.pred3)/(sum(D1$tp.pred3)+sum(D1$fn.pred3))

#Finally, calculate Kappa for your model according to:
kappa2(select(D1, level.up, threshold.pred3))
```
Part III Analysis: 
Threshold 1=0.6125, accuracy=0.826, precision=0.730, recall=0.895, kappa=0.651
Threshold 2=0.8440, accuracy=0.772, precision=0.844, recall=0.528, kappa=0.493
Threshold 3=0.4042, accuracy=0.817, precision=0.702, recall=0.943, kappa=0.639

Threshold 1 was the most accurate with the highest agreement between the predictive model and actual results (kappa). Threshold 2 had the highest precision value at the expense of generating a greater proportion of false negatives (low recall). Threshold 3 had the highest recall value at the expense of generating a greater proportion of false positives (low precision)