---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "12/3/2019"
output: 
  html_document:
    toc: true
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

```{r, result = 'hide'}
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(corrplot)
library(RColorBrewer)
library(rpart)
library(rpart.plot)
library(caret)
```

## Part I

### Upload data
```{r}
tutor <- read.csv("online.data.csv", header = TRUE)
# Recode "level.up": yes = 1, no = 0
tutor_recode <- tutor %>%
      mutate(level.up = ifelse(level.up == "no",0,1))
tutor_recode$level.up <- as.numeric(as.factor(tutor_recode$level.up))
```

### Visualization   
Reference: https://drsimonj.svbtle.com/quick-plot-of-all-variables  
```{r}
# Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
tutor_recode %>%
  keep(is.numeric) %>% 
  gather() %>%   # Convert to key-value pairs
  ggplot(aes(value)) +  # Plot the values
    facet_wrap(~ key, scales = "free") +
    geom_histogram()

# Then visualize the relationships between variables
tutor_corr <- cor(tutor_recode %>% keep(is.numeric))
corrplot(tutor_corr, type="upper", order="hclust",
         col=brewer.pal(n=8, name="RdYlBu"))

# Try to capture an intution about the data and the relationships
# Level up or not is highly correlated with average assignment score, messages sent by each student to other students about the assignments, and post-test score.
```

### Classification tree
```{r}
# create assignment of data
assignment <- sample(1:2, size = nrow(tutor), prob = c(.8, .2), replace = TRUE)

# Create a training and testing data from the original data 
tutor_train <- tutor[assignment == 1, ]    # subset grade to training indices only
tutor_test <- tutor[assignment == 2, ]   # subset grade to test indices only

# Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
tutor.tree <- rpart(formula = level.up ~ av.assignment.score + post.test.score + messages,
                    data = tutor_train,
                    control = rpart.control(minsplit = 1, minbucket = 1))

# Retrieve optimal cp value based on cross-validated error
opt_index <- which.min(tutor.tree$cptable[, "xerror"])
cp_opt <- tutor.tree$cptable[opt_index, "CP"]

# Prune the model (to optimized cp value)
tutor.tree.opt <- prune(tree = tutor.tree, 
                         cp = cp_opt)

# Plot and generate a CP table for your tree 
post(tutor.tree.opt, file = "tutor_tree.ps", title = "Level up") # This creates a pdf image of the tree

# another way to plot
rpart.plot(x = tutor.tree.opt, yesno = 2, type = 0, extra = 0, main = "Level up")

# Generate a probability value that represents the probability that a student levels up based your classification tree 
tutor_test$pred <- predict(tutor.tree.opt, tutor_test, type = "prob")[,2]
# Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
# Intention is to use the label “1” probabilities to compare against actual “level.up” values and come up with ROC curve. We need only the second column (probabilities for label “1”) from the output of “predict” function to be passed as input into “prediction” function.
```



## Part II  
Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.    
```{r}
library(ROCR)

# Plot the curve
pred.detail <- prediction(tutor_test$pred, tutor_test$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

# Calculate the Area Under the Curve
unlist(slot(performance(pred.detail, "auc"), "y.values"))
# Unlist liberates the AUC value from the "performance" object created by ROCR
```

Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?  
The first model is better since AUC is higher than that of the second model.
```{r}
# Create a classification tree using other two variables
tutor.tree2 <- rpart(formula = level.up ~ pre.test.score + forum.posts,
                    data = tutor_train,
                    control = rpart.control(minsplit = 1, minbucket = 1))

# Retrieve optimal cp value based on cross-validated error
opt_index2 <- which.min(tutor.tree2$cptable[, "xerror"])
cp_opt2 <- tutor.tree2$cptable[opt_index2, "CP"]

# Prune the model (to optimized cp value)
tutor.tree.opt2 <- prune(tree = tutor.tree2, 
                         cp = cp_opt2)

# Plot and generate a CP table for your tree 
post(tutor.tree.opt2, file = "tutor_tree2.ps", title = "Level up") # This creates a pdf image of the tree

# another way to plot
rpart.plot(x = tutor.tree.opt2, yesno = 2, type = 0, extra = 0, main = "Level up")

# Generate a probability value that represents the probability that a student levels up based your classification tree 
tutor_test$pred2 <- predict(tutor.tree.opt2, tutor_test, type = "prob")[,2]

# Plot the curve
pred.detail2 <- prediction(tutor_test$pred2, tutor_test$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

# Calculate the Area Under the Curve
unlist(slot(performance(pred.detail2, "auc"), "y.values"))
```



## Part III  
### Thresholds  
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

threshold.pred1 <- 

#Now generate three diagnostics:

D1$accuracy.model1 <-

D1$precision.model1 <- 

D1$recall.model1 <- 

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

```

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.
