---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
data<-read.csv("online.data.csv")
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(tidyverse)
data1<-gather(data,"variable","num")
ggplot(data1, aes(x=num)) + 
  geom_histogram(stat = "count") + 
  facet_wrap(~variable, scales = "free")

#Then visualize the relationships between variables
pairs(data)

#Try to capture an intution about the data and the relationships

# The messages and post.test.score have a extremely strong correlation. messages and
# av.assignment score also has a strong correlation. Therefore, messages might be a 
# good indicator of scores. post.test.scores and w.assignment.score also seems to 
# be correlated.

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
class<-rpart(level.up~av.assignment.score + post.test.score + messages,data)
#Plot and generate a CP table for your tree 
printcp(class)
post(class,file = "class.ps",title = "level.up")
#Generate a probability value that represents the probability that a student levels up based your classification tree 

data$pred <- predict(class, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
library(ROCR)

#Plot the curve
pred.detail <- prediction(data$pred, data$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
class1<-rpart(level.up~pre.test.score + forum.posts,data)
printcp(class1)
post(class1, file = "class1.ps", title = "level up")
data$pred1 <- predict(class1, type = "prob")[,2]
pred.detail1 <- prediction(data$pred1, data$level.up) 
plot(performance(pred.detail1, "tpr", "fpr"))
abline(0, 1, lty = 2)
unlist(slot(performance(pred.detail1,"auc"), "y.values"))
# The first one is a better model. Its cross-validation error is 0, which is smaller than the 0.5875 
# error of the second model. In addition, the AUC value of the first model is larger that of the second
# model. 
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

#Try 0.6 for threshold
threshold.pred1 <- ifelse(data$pred1>0.6,1,0)
data$pred1.result<-ifelse(data$pred1>0.6,1,0)

#calculate the true positive, false positive, true negative and false negative
data$tp<-ifelse(data$level.up=="yes" & data$pred1.result == 1,1,0)
data$fp<-ifelse(data$level.up=="no" & data$pred1.result == 1,1,0)
data$tn<-ifelse(data$level.up=="no" & data$pred1.result == 0,1,0)
data$fn<-ifelse(data$level.up=="yes" & data$pred1.result == 0,1,0)

#Now generate three diagnostics:
data$level.up1<-ifelse(data$level.up=="yes",1,0)
accuracy.model1 <- mean(data$pred1.result == data$level.up1)

precision.model1 <- sum(data$tp)/(sum(data$tp)+sum(data$fp))

recall.model1 <- sum(data$tp)/(sum(data$tp)+sum(data$fn))

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(data$level.up1, data$pred1.result)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)

#I get a kappa value of 1.05 which is more than 1. I will try another way I found online
#observed accuracy
oa<-mean(data$pred1.result == data$level.up1)
#expected accuracy
tp<-sum(data$tp)
fp<-sum(data$fp)
fn<-sum(data$fn)
tn<-sum(data$tn)
ea<-((tp+fp)/1000)*((tp+fn)/1000)+((tn+fn)/1000)*((tn+fp)/1000)
#kappa=(oa-ea)/(1-ea)
k<-(oa-ea)/(1-ea)

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
#Try 0.3 for threshold
threshold.pred2 <- ifelse(data$pred1>0.3,1,0)
data$pred2.result<-ifelse(data$pred1>0.3,1,0)

#calculate the true positive, false positive, true negative and false negative
data$tp1<-ifelse(data$level.up=="yes" & data$pred2.result == 1,1,0)
data$fp1<-ifelse(data$level.up=="no" & data$pred2.result == 1,1,0)
data$tn1<-ifelse(data$level.up=="no" & data$pred2.result == 0,1,0)
data$fn1<-ifelse(data$level.up=="yes" & data$pred2.result == 0,1,0)

#Now generate three diagnostics:
accuracy.model2 <- mean(data$pred2.result == data$level.up1)

precision.model2 <- sum(data$tp1)/(sum(data$tp1)+sum(data$fp1))

recall.model2 <- sum(data$tp1)/(sum(data$tp1)+sum(data$fn1))

#Finally, calculate Kappa for your model according to:
#observed accuracy
oa1<-mean(data$pred2.result == data$level.up1)
#expected accuracy
tp1<-sum(data$tp1)
fp1<-sum(data$fp1)
fn1<-sum(data$fn1)
tn1<-sum(data$tn1)
ea1<-((tp1+fp1)/1000)*((tp1+fn1)/1000)+((tn1+fn1)/1000)*((tn1+fp1)/1000)
#kappa=(oa-ea)/(1-ea)
k1<-(oa1-ea1)/(1-ea1)






```
model 1
Threshold - 0.6
kappa - 0.557
accuracy - 0.783
precision - 0.705
recall - 0.788

model 2
Threshold - 0.3
kappa - 0.547
accuracy - 0.769
precision - 0.654
recall - 0.895

The model with 0.6 threshold has slightly higher kappa and accuracy compares to the model with 0.3 threhold. That indicates model with 0.6 threhold will have a better prediction. I didn't try 0.5 threshold, but I think it's worth mentioning that the reason of 0.6 threshold has a better prediction might be it's closer to 0.5 compares to 0.3 threhold. 

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then commit, push and pull request both the RMarkdown file and the html file.
