---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
D1<-read.csv("online.data.csv")
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(ggplot2)
M1<-ggplot(D1, aes(y=pre.test.score)) + geom_histogram()
M1
M2<-ggplot(D1, aes(x=post.test.score)) + geom_histogram(colour="white")
M3<-ggplot(D1, aes(x=pre.test.score)) + geom_histogram(colour="white")
M4<-ggplot(D1, aes(x=messages)) + geom_histogram(colour="white")
M5<-ggplot(D1, aes(x=forum.posts)) + geom_histogram(colour="white")
M6<-ggplot(D1, aes(x=av.assignment.score)) + geom_histogram(colour="white")
M7<-ggplot(D1, aes(x=level.up)) + geom_histogram(colour="white")

#Then visualize the relationships between variables
C1<-ggplot(D1, aes(x=pre.test.score, y=post.test.score)) + geom_point(shape=1)
C1 + facet_grid(level.up ~ .)
C2<-ggplot(D1, aes(x=id,y=pre.test.score)) + geom_point(shape=1)
C2 + facet_grid(level.up ~ .)
C3<-ggplot(D1, aes(x=id,y=post.test.score)) + geom_point(shape=1)
C3 + facet_grid(level.up ~ .)
C4<-ggplot(D1, aes(x=messages, y=post.test.score)) + geom_point(shape=1)
C4+facet_grid(level.up ~ .) #interesting graph shows students who level up usually posted more messages
C5<-ggplot(D1,aes(x=forum.posts, y=post.test.score)) + geom_point()
C5+facet_grid(level.up ~ .) #doesn't show any obvious correlation between forum.posts and post.test.score
C6<-ggplot(D1,aes(x=av.assignment.score, y=post.test.score)) + geom_point()
C6+facet_grid(level.up ~ .) # shows that those students who leveled up usually scored higher on the assignments

#Try to capture an intution about the data and the relationships

#The students who posted > 90 messages usually leveled up. Then if a student had an average assignment score of > 0.25 then they usually leveled up. The students who scored > 0.5 on the pre.test.score usually leveled up as well. Forum.posts did not show an obvious pattern.
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

#The three variables I will use is pre.test.score, messages, and av.assignment.score.
library(rpart)

c.tree1 <- rpart(level.up ~ pre.test.score + messages + av.assignment.score , method="class", data=D1)

printcp(c.tree1)

post(c.tree1, file = "ass7_tree1.ps", title = "Online Geography Tutoring System")


#Plot and generate a CP table for your tree 

D1$predict1 <- predict(c.tree1, D1, type = "class")

table(D1$level.up, D1$predict1)

#Generate a probability value that represents the probability that a student levels up based your classification tree 

#Using the equation for Cohen's Kappa, p0 = (390+595)/ 1000 = .985, pe = (605/1000)*(600/1000)+(400/1000)*(395/1000) = 0.521, Kappa = (.985-.521)/(1-.521) = 0.9687

#If this is right then I have a pretty good model! My Cohen Kappa is .9687 so the agreement between my model and the data is pretty good. Is it too good? Is it overfit? Ummm...


D1$pred <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
install.packages("ROCR")
library(ROCR)

#Plot the curve
pred.detail <- prediction(D1$pred, D1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values")) #Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

#For this part I will use the post.test.score and forum.posts variables.
c.tree2 <- rpart(level.up ~ post.test.score +forum.posts , method="class", data=D1)
D1$pred2 <- predict(c.tree2, type = "prob")[,2]

pred2.detail <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred2.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred2.detail,"auc"), "y.values"))

#My model using messages, av.assignment.score, and pre.test.score had a higher AUC than this model with post.test.score and forum.posts. Thus, my model was better.
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

D1$threshold.pred1 <- 0.001 #just sort of arbitrarily chosen as a very close value to the upper left corner of the ROC curve and slightly lower than the AUC value

#Now generate three diagnostics:

#accuracy = correct predictions/ total predictions
D1$accuracy.model1 <- (390+595)/(1000)

#Precision = TP/(TP+FP)
D1$precision.model1 <- 390/(390+5)

#Recall = TP/(TP+FN)
D1$recall.model1 <- 390/(390+10)

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

#I turned my threshold value all the way down to 0.01 and I was still getting a Kappa of "1". Is that first rise of my ROC, the one that is almost vertical, really that steep? I don't know why my threshold reading is off.

```


