---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
C1 <- read.csv("online.data.csv")

```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)

library(reshape2)
library(plotly)
library(tidyr)
#data is in wide and needs to be in long, variable names need to be on top of eachother in column

#take out ID first, to remove a column we can use subset and a - befor C(<columnname>)
C2 <- subset(C1,select=-c(id))

C2$level.up <- ifelse(C2$level.up == "no",0,1)
View(C2)

#need to scale the data, but make sure to set it up as.data.frame
C3 <- as.data.frame(scale(C2))
View(C3)

#for the facet to work, data needs to be in "long" form. Gather function moves columns to new row, tidyr::gather(data,variablecolumnname,datacolumnname,whichcolumnsfromoriginaltopull)
C4 <- tidyr::gather(C3,variable,value,1:6) 

View(C4)

#facet
library(ggplot2)

hist1 = ggplot(C4,aes(x=value)) + geom_histogram(binwidth = 0.1) + facet_wrap(~variable, scales = "free")
hist1

#Then visualize the relationships between variables
library(corrplot)

pairs(C3)

cor1 <- cor(C3)

corrplot(cor1, order="AOE", method="circle", tl.pos="lt", type="upper", tl.col="black", tl.cex=0.6, tl.srt=45, addCoef.col="black", addCoefasPercent = TRUE, sig.level=0.50, insig = "blank")

#Try to capture an intution about the data and the relationships
#The strongest correlation is between messages and post test score. Average assignment score has a high correlation with level up

```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
c.tree1 <- rpart(level.up ~ messages + forum.posts + av.assignment.score, method="class" , data = C2, control=rpart.control(minsplit=1,minbucket=1,cp=0.001))

printcp(c.tree1)
#Plot and generate a CP table for your tree 

plot(c.tree1)
text(c.tree1)

#Generate a probability value that represents the probability that a student levels up based your classification tree 

C2$pred <- predict(c.tree1, type = "prob")[,2]
View(C2)
#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail <- prediction(C2$pred, C2$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
pred <- prediction(C2$pred, C2$level.up)
unlist(slot(performance(pred,"auc"), "y.values"))
# Area Under Curve = 0.9959375

#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model 
c.tree2 <- rpart(level.up ~ pre.test.score + post.test.score, method="class" , data = C2, control=rpart.control(minsplit=1,minbucket=1,cp=0.001))

printcp(c.tree2)
#Plot and generate a CP table for your tree 

plot(c.tree2)
text(c.tree2)

C2$pred2 <- predict(c.tree2, type = "prob")[,2]
View(C2)

pred.detail2 <- prediction(C2$pred2, C2$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

pred2 <- prediction(C2$pred2, C2$level.up)
unlist(slot(performance(pred2,"auc"), "y.values"))
#Area Under Curve .9740771

#compare the plots & results of your two models. Which one do you think was the better model? Why?
#Both results seem pretty good, though the area under the curve is larger (.9959375) when predicting using messages, forum posts, and avg.assignmentscore, than when using just pretest and posttest scores (.9740771).

```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

C2$threshold.pred1 <- ifelse(C2$pred>.65, 1, 0)
View(C2)

tableA <- table(C2$level.up, C2$threshold.pred1) 
tableA

#      0   1
#  0 595   5
#  1   0 400

#Now generate three diagnostics:

C2$accuracy.model1 <- (595+400)/(595+400+5) #0.995

C2$precision.model1 <- (400)/(400+0) #1.00

C2$recall.model1 <- (400)/(400+5) #.9876543

#Finally, calculate Kappa for your model according to:

#Convert to matrix
matrix1 <- as.matrix(tableA)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1)

#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

C2$threshold2.pred <- ifelse(C2$pred>.90, 1, 0)
View(C2)

tableB <- table(C2$level.up, C2$threshold2.pred) 
tableB
#TableB and TableA are the same with a threshold of .65 and .9, so the diagnostics are going to be the same. This is a result of the probabilities being either 1, 0, or .9874

#I'm going to try this on my second model instead, because there is more variation in the probabilities

C2$threshold1.pred2 <- ifelse(C2$pred2>.65, 1, 0)
View(C2)

tableC <- table(C2$level.up, C2$threshold1.pred2) 
tableC

#      0   1
#  0 555  45
#  1  27 373

C2$accuracy.model2 <- (555+373)/(555+45+27+373) #0.928

C2$precision.model2 <- (373)/(373+27) #0.9325

C2$recall.model2 <- (373)/(373+45) #0.8923


#Threshold of .9 on the second model
C2$threshold2.pred2 <- ifelse(C2$pred2>.9, 1, 0)
View(C2)

tableD <- table(C2$level.up, C2$threshold2.pred2) 
tableD

#    0   1
#  0 596   4
#  1 187 213

C2$accuracy2.model2 <- (596+213)/(596+4+187+213) #0.809

C2$precision2.model2 <- (213)/(213+187) #0.5325

C2$recall2.model2 <- (213)/(213+4) #0.9815

#As we change the threshold on the second model to become higher (from .65 to .9), the accuracy of the model decreases. The precision of the model also decreases, as there are more false positives. But, the recall increases, as there is a lower percentage of false negatives.

matrix2 <- as.matrix(tableD)
kappa(matrix2, exact = TRUE)/kappa(matrix2)

#Again I get a kappa of >1, based on my reading (aka wikipedia) these kappa values should be between 0 and 1. Not quite sure what's going on here....
```


