---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
D1 <- read.csv("online.data.csv")
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(ggplot2)
library(dplyr)
library(tidyr)
D2 <- select(D1, 1:7)
D2$level.up <- ifelse(D1$level.up == "yes", 1,0)
D3 <- gather(D2, "measure", "score", 2:7)
p1 <- ggplot(D3, aes(score)) + facet_wrap(~measure, scales = "free")
p1 + geom_histogram(stat = "count")
#Then visualize the relationships between variables
pairs(D2)
#Try to capture an intution about the data and the relationships
#Strong positive correlation between post.test.score and messages, which simulates an almost linear relationship in the graph. Therefore, more messages sent by the student might indicate a higher post test score. 
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
c.tree1 <- rpart(level.up ~ av.assignment.score + messages + post.test.score, method="class", data=D1)
#Plot and generate a CP table for your tree 
printcp(c.tree1)
plot(c.tree1)
text(c.tree1)
#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
library(ROCR)

#Plot the curve
pred.detail <- prediction(D1$pred, D1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR
#AUC = 1

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
c.tree2 <- rpart(level.up ~ forum.posts + pre.test.score, method="class", data=D1)
D1$pred2 <- predict(c.tree2, type = "prob")[,2]
pred.detail2 <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)
unlist(slot(performance(pred.detail2,"auc"), "y.values"))
#AUC = 0.818
#The model consists of av.assignment.score, messages, and post.test.score has a AUC value of 1, which means that it's better at predicting the level-up system. 
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.
D1$threshold.pred1 <- ifelse(D1$pred >= 0.95, "yes", "no")

#Now generate three diagnostics:

accuracy.model1 <- mean(ifelse(D1$level.up == D1$threshold.pred1, 1, 0)) #1

D1$truepos.model1 <- ifelse(D1$level.up == "yes" & D1$threshold.pred1 == "yes", 1, 0)
D1$falsepos.model1 <- ifelse(D1$level.up == "no" & D1$threshold.pred1 == "yes", 1,0)
D1$falseneg.model1 <- ifelse(D1$level.up == "yes" & D1$threshold.pred1 == "no", 1,0)

precision.model1 <- sum(D1$truepos.model1)/(sum(D1$truepos.model1) + sum(D1$falsepos.model1)) #1

recall.model1 <- sum(D1$truepos.model1)/(sum(D1$truepos.model1) + sum(D1$falseneg.model1)) #1

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
kappa(matrix1, exact = TRUE)/kappa(matrix1) #kappa value = 1.154


#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
D1$threshold.pred1a <- ifelse(D1$pred >= 0.6, "yes", "no")
accuracy.model1a <- mean(ifelse(D1$level.up == D1$threshold.pred1a, 1, 0)) #1
D1$truepos.model1a <- ifelse(D1$level.up == "yes" & D1$threshold.pred1a == "yes", 1, 0)
D1$falsepos.model1a <- ifelse(D1$level.up == "no" & D1$threshold.pred1a == "yes", 1,0)
D1$falseneg.model1a <- ifelse(D1$level.up == "yes" & D1$threshold.pred1a == "no", 1,0)
precision.model1a <- sum(D1$truepos.model1a)/(sum(D1$truepos.model1a) + sum(D1$falsepos.model1a)) #1
recall.model1a <- sum(D1$truepos.model1a)/(sum(D1$truepos.model1a) + sum(D1$falseneg.model1a)) #1
table1a <- table(D1$level.up, D1$threshold.pred1a)
matrix1a <- as.matrix(table1a)
kappa(matrix1a, exact = TRUE)/kappa(matrix1a) #kappa value = 1.154

D1$threshold.pred1b <- ifelse(D1$pred >= 0.25, "yes", "no")
accuracy.model1b <- mean(ifelse(D1$level.up == D1$threshold.pred1b, 1, 0)) #1
D1$truepos.model1b <- ifelse(D1$level.up == "yes" & D1$threshold.pred1b == "yes", 1, 0)
D1$falsepos.model1b <- ifelse(D1$level.up == "no" & D1$threshold.pred1b == "yes", 1,0)
D1$falseneg.model1b <- ifelse(D1$level.up == "yes" & D1$threshold.pred1b == "no", 1,0)
precision.model1b <- sum(D1$truepos.model1b)/(sum(D1$truepos.model1b) + sum(D1$falsepos.model1b)) #1
recall.model1b <- sum(D1$truepos.model1b)/(sum(D1$truepos.model1b) + sum(D1$falseneg.model1b)) #1
table1b <- table(D1$level.up, D1$threshold.pred1b)
matrix1b <- as.matrix(table1b)
kappa(matrix1b, exact = TRUE)/kappa(matrix1b) #kappa value = 1.154

#For model 1, changes of threshold yielded the same accuracy, precision, recall and kappa values. 

#For model 2
D1$threshold.pred2 <- ifelse(D1$pred2 >= 0.25, "yes", "no")
accuracy.model2 <- mean(ifelse(D1$level.up == D1$threshold.pred2, 1, 0)) #0.769
D1$truepos.model2 <- ifelse(D1$level.up == "yes" & D1$threshold.pred2 == "yes", 1, 0)
D1$falsepos.model2 <- ifelse(D1$level.up == "no" & D1$threshold.pred2 == "yes", 1,0)
D1$falseneg.model2 <- ifelse(D1$level.up == "yes" & D1$threshold.pred2 == "no", 1,0)
precision.model2 <- sum(D1$truepos.model2)/(sum(D1$truepos.model2) + sum(D1$falsepos.model2)) #0.654
recall.model2 <- sum(D1$truepos.model2)/(sum(D1$truepos.model2) + sum(D1$falseneg.model2)) #0.895
table2 <- table(D1$level.up, D1$threshold.pred2)
matrix2 <- as.matrix(table2)
kappa(matrix2, exact = TRUE)/kappa(matrix2) #kappa value = 0.868

D1$threshold.pred2a <- ifelse(D1$pred2 >= 0.6, "yes", "no")
accuracy.model2a <- mean(ifelse(D1$level.up == D1$threshold.pred2a, 1, 0)) #0.783
D1$truepos.model2a <- ifelse(D1$level.up == "yes" & D1$threshold.pred2a == "yes", 1, 0)
D1$falsepos.model2a <- ifelse(D1$level.up == "no" & D1$threshold.pred2a == "yes", 1,0)
D1$falseneg.model2a <- ifelse(D1$level.up == "yes" & D1$threshold.pred2a == "no", 1,0)
precision.model2a <- sum(D1$truepos.model2a)/(sum(D1$truepos.model2a) + sum(D1$falsepos.model2a)) #0.705
recall.model2a <- sum(D1$truepos.model2a)/(sum(D1$truepos.model2a) + sum(D1$falseneg.model2a)) #0.788
table2a <- table(D1$level.up, D1$threshold.pred2a)
matrix2a <- as.matrix(table2a)
kappa(matrix2a, exact = TRUE)/kappa(matrix2a) #kappa value = 1.056
#Threshold 2 generates higher accuracy, precision, and kappa values at the cost of lower recall.
```


