---
title: "Assignment 7"
author: "Oren Ross"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}

GAME1 = read.csv("online.data.csv", header = TRUE)
library(ggplot2)
library(rpart)
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)


ggplot(GAME1, aes(pre.test.score)) + geom_histogram()
ggplot(GAME1, aes(post.test.score)) + geom_histogram()
ggplot(GAME1, aes(messages)) + geom_histogram()
ggplot(GAME1, aes(forum.posts)) + geom_histogram()
ggplot(GAME1, aes(av.assignment.score)) + geom_histogram()


ggplot(GAME1, aes(pre.test.score, post.test.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(messages, forum.posts)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(messages, av.assignment.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(forum.posts, av.assignment.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(messages, post.test.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(forum.posts, post.test.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(av.assignment.score, post.test.score)) + geom_point() + facet_grid(. ~ level.up)




#Then visualize the relationships between variables


ggplot(GAME1, aes(pre.test.score, post.test.score)) + geom_point() + facet_grid(level.up ~ .)

ggplot(GAME1, aes(pre.test.score, post.test.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(messages, forum.posts)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(messages, av.assignment.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(forum.posts, av.assignment.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(messages, post.test.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(forum.posts, post.test.score)) + geom_point() + facet_grid(. ~ level.up)

ggplot(GAME1, aes(av.assignment.score, post.test.score)) + geom_point() + facet_grid(. ~ level.up)


#Try to capture an intution about the data and the relationships

#Based on the graphs, it seems that the AV score uses the messages and forum posts, which is positively correlated with post test score, even though the pre test scores are higher for those that leveled up.


```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)

# 3 variables - post.test.score, av.assigment.score, messages


GAME_Tree1 <- rpart(level.up ~ post.test.score + messages + av.assignment.score, method = "class", data=GAME1)

printcp(GAME_Tree1)

post(GAME_Tree1, file = "Game_Tree1.ps", title = "Game") #This creates a pdf image of the tree

#Plot and generate a CP table for your tree 

GAME_Tree2 <- prune(GAME_Tree1, cp = 0.1)
printcp(GAME_Tree2)
post(GAME_Tree2, file = "Game_Tree2.ps", title = "Game")


GAME_Tree3 <- prune(GAME_Tree1, cp = 0.000001)
printcp(GAME_Tree3)
post(GAME_Tree3, file = "Game_Tree3.ps", title = "Game")
#Tested out some different CP levels, and it seems like the original is best. No need to prune.

#Generate a probability value that represents the probability that a student levels up based your classification tree 

GAME1$pred <- predict(GAME_Tree1, type = "prob")[,2]

#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

```

## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

```{r}

#install.packages('ROCR')
#library(ROCR)

#Plot the curve
pred.detail <- prediction(GAME1$pred, GAME1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))

#Unlist liberates the AUC value from the "performance" object created by ROCR


```


```{r}

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

# 3 variables - pre.test.score, forum.posts


GAME_Tree2.1 <- rpart(level.up ~ pre.test.score + forum.posts, method = "class", data=GAME1)

printcp(GAME_Tree2.1)

post(GAME_Tree2.1, file = "Game_Tree2.1.ps", title = "Game") #This creates a pdf image of the tree

#Plot and generate a CP table for your tree 

GAME_Tree2.2 <- prune(GAME_Tree2.1, cp = 0.1)
printcp(GAME_Tree2.2)
post(GAME_Tree2.2, file = "Game_Tree2.2.ps", title = "Game")


GAME_Tree2.3 <- prune(GAME_Tree2.1, cp = 0.000001)
printcp(GAME_Tree2.3)
post(GAME_Tree2.3, file = "Game_Tree2.3.ps", title = "Game")
#Tested out some different CP levels, and it seems like the original is best. No need to prune.

#Generate a probability value that represents the probability that a student levels up based your classification tree 

GAME1$pred2 <- predict(GAME_Tree2.1, type = "prob")[,2]

#Plot the curve
pred.detail2 <- prediction(GAME1$pred2, GAME1$level.up) 
plot(performance(pred.detail2, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail2,"auc"), "y.values"))

#The first model was best, since it basically was a perfect predictor. The second model had the cross validate error rate of .5875 which is not very good. It also has an AUC of .818. This means the first model does a much better job prediciting than the second. In the end, only two variables are really needed for good predictions, av.assignment.score and post.scores.

```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

#Since my first model is actually "perfect", I have decided to use my 2nd model. This should make for a more interesting test

GAME1$threshold.pred1 <- ifelse(GAME1$pred2 > 0.5, 1, 0)
table(GAME1$threshold.pred1, GAME1$level.up)

#Since the model is not very good, I am starting with .5, so I can make the next model a little better.

#Now generate three diagnostics:


GAME1$accuracy.model1 <- (sum((GAME1$level.up == "no")&(GAME1$threshold.pred1 == "0")) + sum((GAME1$level.up == "yes")&(GAME1$threshold.pred1 == 1))) / nrow(GAME1)


GAME1$precision.model1 <- sum((GAME1$level.up == "yes")&(GAME1$threshold.pred1 == 1)) / (sum((GAME1$level.up == "yes")&(GAME1$threshold.pred1 == 0)) + sum((GAME1$level.up == "yes")&(GAME1$threshold.pred1 == 1)))



GAME1$recall.model1 <- (sum((GAME1$level.up == "yes")&(GAME1$threshold.pred1 == 1)) / (sum((GAME1$level.up == "yes")&(GAME1$threshold.pred1 == 1)) + sum((GAME1$level.up == "no")&(GAME1$threshold.pred1 == 1))))



#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(GAME1$level.up, GAME1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa


kappa(matrix1, exact = TRUE)/kappa(matrix1)

#Apparently, you need to load irr and use the following code to get this to work.
library(irr)
kappa2(GAME1[,c(8,10)], "unweighted")



#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

#I upped the threshold to .7

GAME1$threshold.pred2 <- ifelse(GAME1$pred2 > 0.7, 1, 0)
table(GAME1$threshold.pred2, GAME1$level.up)

#Since the model is not very good, I am starting with .5, so I can make the next model a little better.

#Now generate three diagnostics:




GAME1$accuracy.model2 <- (sum((GAME1$level.up == "no")&(GAME1$threshold.pred2 == "0")) + sum((GAME1$level.up == "yes")&(GAME1$threshold.pred2 == 1))) / nrow(GAME1)


GAME1$precision.model2 <- sum((GAME1$level.up == "yes")&(GAME1$threshold.pred2 == 1)) / (sum((GAME1$level.up == "yes")&(GAME1$threshold.pred2 == 0)) + sum((GAME1$level.up == "yes")&(GAME1$threshold.pred2 == 1)))



GAME1$recall.model2 <- (sum((GAME1$level.up == "yes")&(GAME1$threshold.pred2 == 1)) / (sum((GAME1$level.up == "yes")&(GAME1$threshold.pred2 == 1)) + sum((GAME1$level.up == "no")&(GAME1$threshold.pred2 == 1))))


#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table2 <- table(GAME1$level.up, GAME1$threshold.pred2)

#Convert to matrix
matrix2 <- as.matrix(table2)

#Calculate kappa


kappa(matrix2, exact = TRUE)/kappa(matrix2)

#Apparently, you need to load irr and use the following code to get this to work.
library(irr)
kappa2(GAME1[,c(8,12)], "unweighted")

#There is a minimal difference, but it is not very good. It looks like my second model is very bad compared to my perfect 1

```


