---
title: "Assignment 7 - Answers"
author: "Charles Lang"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

## Part I

#Upload data
```{r}
A1<-read.csv("online.data.csv",header=T)
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library(dplyr)
library(tidyr)
library(ggplot2)
A2 <- dplyr::select(A1, -1) #omit Column "id"
A2$level.up <- ifelse(A2$level.up=="yes", 1, 0) #convert yes/no into 1/0
A3 <- tidyr::gather(A2, "variables", "value", 1:6) #convert the wide version into the long version
library(ggplot2)
 ggplot(A3, aes(value))+
    geom_histogram(data = A3[A3$variables=="post.test.score",], binwidth = 0.01)+
    geom_histogram(data = A3[A3$variables=="pre.test.score",], binwidth = 0.01)+
    geom_histogram(data = A3[A3$variables=="messages",], binwidth = 5)+
    geom_histogram(data = A3[A3$variables=="forum.posts",], binwidth = 1)+geom_histogram(data = A3[A3$variables=="av.assignment.score",], binwidth = 0.01)+
    geom_histogram(data = A3[A3$variables=="level.up",], binwidth = 0.1)+
    facet_wrap(~variables, scale="free")
#Then visualize the relationships between variables
plot(A2) #the scatter plot between variables
COR <- cor(A2) # generate pairwise correlations
 library(corrplot)
 corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
 tl.col="black", tl.cex=0.6, tl.srt=45, 
         addCoef.col="black", addCoefasPercent = TRUE,
         sig.level=0.50, insig = "blank") #the correlation plot between variables
#Try to capture an intution about the data and the relationships
#variable messages has an unexpectedly high correlation with variable post.test.score. However,variable forum. posts doesn't highly correlate with other variables. It seems that direct communication between 5th grade students (e.g. messaging) is a more effective way than indirect communication (e.g. forum posting) to improve their grades of geography. 
```
#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library(rpart)
 c.tree1 <- rpart(level.up ~ messages + post.test.score + av.assignment.score, method="class", data=A2, control = rpart.control(minsplit=1, minbucket=1, cp=0.01))
 #Plot and generate a CP table for your tree 

 printcp(c.tree1)
 post(c.tree1, file = "tree.ps", title = "predicting levels up")
#Generate a probability value that represents the probability that a student levels up based your classification tree 
A2$pred <- predict(c.tree1, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.
```
## Part II
#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.
```{r}
install.packages("ROCR")
library(ROCR)

#Plot the curve
pred.detail <- prediction(A2$pred, A2$level.up) 
 plot(performance(pred.detail, "tpr", "fpr"))
 abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR

#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?
c.tree2 <- rpart(level.up ~ pre.test.score + forum.posts, method="class", data=A2, control = rpart.control(minsplit=1, minbucket=1))
 printcp(c.tree2)
 post(c.tree2, file = "tree2.ps", title = "predicting levels up")

 A2$pred2 <- predict(c.tree2, type = "prob")[,2]
 pred.detail2 <- prediction(A2$pred2, A2$level.up) 
 plot(performance(pred.detail2, "tpr", "fpr"))
 abline(0, 1, lty = 2)
unlist(slot(performance(pred.detail2,"auc"), "y.values"))
```
## Part III
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.

A2$threshold.pred1 <- ifelse(A2$pred2>0.7, 1, 0) 
#Now generate three diagnostics:
A4.1 <- dplyr::count(A2, level.up, threshold.pred1)
 A4.1$type <- ifelse(A4.1$level.up==1 & A4.1$threshold.pred1==1, "TP", ifelse(A4.1$level.up==1 & A4.1$threshold.pred1==0, "FN", ifelse(A4.1$level.up==0 & A4.1$threshold.pred1==1, "FP", "TN")))
 A5 <- data.frame((A4.1[A4.1$type=="TP",]$n + A4.1[A4.1$type=="TN",]$n) / sum(A4.1$n)) #calculate accuracy
 names(A5) <- c("accuracy.model1")
 A5$precision.model1 <- A4.1[A4.1$type=="TP",]$n / (A4.1[A4.1$type=="TP",]$n + A4.1[A4.1$type=="FP",]$n) #calculate precision
 A5$recall.model1 <- A4.1[A4.1$type=="TP",]$n / (A4.1[A4.1$type=="TP",]$n + A4.1[A4.1$type=="TN",]$n) #calculate recall

#Finally, calculate Kappa for your model according to:

 #First generate the table of comparisons
 table1 <- table(A2$level.up, A2$threshold.pred1)

 #Convert to matrix
 matrix1 <- as.matrix(table1)

 #Calculate kappa
 A5$kappa.model1 <- kappa(matrix1, exact = TRUE)/kappa(matrix1)

 #Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?

 A2$threshold.pred2 <- ifelse(A2$pred2>0.6, 1, 0) #a new threshold 0.6

 A4.2 <- dplyr::count(A2, level.up, threshold.pred2)
 A4.2$type <- ifelse(A4.2$level.up==1 & A4.2$threshold.pred2==1, "TP", ifelse(A4.2$level.up==1 & A4.2$threshold.pred2==0, "FN", ifelse(A4.2$level.up==0 & A4.2$threshold.pred2==1, "FP", "TN")))

 A5$accuracy.model2 <- (A4.2[A4.2$type=="TP",]$n + A4.2[A4.2$type=="TN",]$n) / sum(A4.2$n)

 A5$precision.model2 <- A4.2[A4.2$type=="TP",]$n / (A4.2[A4.2$type=="TP",]$n + A4.2[A4.2$type=="FP",]$n)

 A5$recall.model2 <- A4.2[A4.2$type=="TP",]$n / (A4.2[A4.2$type=="TP",]$n + A4.2[A4.2$type=="TN",]$n)

 #New Kappa
 table2 <- table(A2$level.up, A2$threshold.pred2)
 matrix2 <- as.matrix(table2)
 A5$kappa.model2 <- kappa(matrix2, exact = TRUE)/kappa(matrix2)


